---
always_allow_html: true
---

# Archaeobotany {#sec-archaeobotany style="text-align:justify;"}

```{r}
#| echo: false

# Load functions
source('functions_bot.R')
```

::: callout-important
## Page under construction {style="text-align:justify;"}

The results presented here are preliminary and the chapter has yet to be written.
:::

In this chapter, I will present the macrobotanical data from 190 case studies used to carry on this research (@sec-materials-bot), along with the statistics performed on the data. The results will be first presented temporally, and a discussion of the diachronic trends will follow at the end of the chapter.

## Case studies {style="text-align:justify;"}

The following map shows the sites under investigation, divided by chronology. Please select the desired chronology (or chronologies) from the legend at the top-right of the map.

```{r}
#| echo: false
#| output: false

# Load libraries
library(sp)
library(rgdal)
library(leaflet)
library(leaflet.extras)
library(htmltools)
library(tidyverse)
library(fontawesome) 

# Get an Italy shapefile
#download.file(url = 'http://biogeo.ucdavis.edu/data/diva/adm/ITA_adm.zip', 
            #  destfile = 'italy.zip')
#unzip(zipfile = 'italy.zip')

italy <- readOGR('Italy_SHP/ITA_Peninsula.shp')
bot_map <- read.csv("/Users/robertoragno/Desktop/University/Bari/PhD - Quarto/Database export/plants.csv", header=TRUE, sep=";")

# Grouping
bot_map_R <- filter(bot_map, Chronology=="R")
bot_map_LR<- filter(bot_map, Chronology=="LR")
bot_map_EMA<- filter(bot_map, Chronology=="EMA")
bot_map_Ma<- filter(bot_map, Chronology=="Ma")

#Icon for the map

Icon_red <- awesomeIcons(
  text = fa("wheat-awn"),
  iconColor = 'yellow',
  library = 'fa',
  markerColor = "red"
)

Icon_lightred <- awesomeIcons(
  text = fa("wheat-awn"),
  iconColor = 'yellow',
  library = 'glyphicon',
  markerColor = "lightred"
)

Icon_lightgray <- awesomeIcons(
  text = fa("wheat-awn"),
  iconColor = 'black',
  library = 'glyphicon',
  markerColor = "lightgray"
)

Icon_green <- awesomeIcons(
  text = fa("wheat-awn"),
  iconColor = 'darkgreen',
  library = 'glyphicon',
  markerColor = "lightgreen"
)

# Build the map
italy_bot_map <- leaflet(data=italy) %>%
  addProviderTiles(providers$CartoDB.Positron,  group = "Political") %>%
  addProviderTiles(providers$Esri.WorldPhysical, group = "Physical") %>%  addPolygons(weight = 1, 
              smoothFactor = 0.5,
              opacity = 0.2, 
              fillOpacity = 0.5,
              fillColor = "#F9F5EB",
              color="#D7A86E"
              ) %>%
  addAwesomeMarkers(data=bot_map_R, lng = bot_map_R$x, lat = bot_map_R$y,
                    group="R", 
                    icon = Icon_red,
                    popup = paste(
                      "<b>", bot_map_R$site_name, "</b>",
                      "<hr style='border:2px solid green; color: solid green, border-radius:2px; margin-top:1.5px; margin-bottom:5px'>",
                   "<b>Site ID:</b>", bot_map_R$site_code,"<br>",
                           "<b>Type:</b>", bot_map_R$type_name, "<br>",
                           "<b>Centuries:</b>", bot_map_R$data_valid_start,
                           "-", bot_map_R$data_valid_end, "<br>",
                           "<b>Chronology:</b>", bot_map_R$Chronology, "<br>",
                           "<b>Reference:</b><i>", bot_map_R$short_ref, "</i>"
                           ), 
             label=~htmlEscape(bot_map_R$site_name)
 ) %>%
    addAwesomeMarkers(data=bot_map_LR, lng = bot_map_LR$x, lat = bot_map_LR$y,
                    group="LR",
                    icon = Icon_lightred,
                    popup = paste(
                      "<b>", bot_map_LR$site_name, "</b>",
                      "<hr style='border:2px solid green; color: solid green, border-radius:2px; margin-top:1.5px; margin-bottom:5px'>",
                   "<b>Site ID:</b>", bot_map_LR$site_code,"<br>",
                           "<b>Type:</b>", bot_map_LR$type_name, "<br>",
                           "<b>Centuries:</b>", bot_map_LR$data_valid_start,
                           "-", bot_map_LR$data_valid_end, "<br>",
                           "<b>Chronology:</b>", bot_map_LR$Chronology, "<br>",
                           "<b>Reference:</b><i>", bot_map_LR$short_ref, "</i>"
                           ), 
             label=~htmlEscape(bot_map_LR$site_name)
 ) %>%
      addAwesomeMarkers(data=bot_map_EMA, lng = bot_map_EMA$x, lat = bot_map_EMA$y,
                    group="EMA",
                    icon = Icon_lightgray,
                    popup = paste(
                      "<b>", bot_map_EMA$site_name, "</b>",
                      "<hr style='border:2px solid green; color: solid green, border-radius:2px; margin-top:1.5px; margin-bottom:5px'>",
                   "<b>Site ID:</b>", bot_map_EMA$site_code,"<br>",
                           "<b>Type:</b>", bot_map_EMA$type_name, "<br>",
                           "<b>Centuries:</b>", bot_map_EMA$data_valid_start,
                           "-", bot_map_EMA$data_valid_end, "<br>",
                           "<b>Chronology:</b>", bot_map_EMA$Chronology, "<br>",
                           "<b>Reference:</b><i>", bot_map_EMA$short_ref, "</i>"
                           ), 
             label=~htmlEscape(bot_map_EMA$site_name)
 ) %>%
        addAwesomeMarkers(data=bot_map_Ma, lng = bot_map_Ma$x, lat = bot_map_Ma$y,
                    group="Ma",
                    icon = Icon_green,
                    popup = paste(
                      "<b>", bot_map_Ma$site_name, "</b>",
                      "<hr style='border:2px solid green; color: solid green, border-radius:2px; margin-top:1.5px; margin-bottom:5px'>",
                   "<b>Site ID:</b>", bot_map_Ma$site_code,"<br>",
                           "<b>Type:</b>", bot_map_Ma$type_name, "<br>",
                           "<b>Centuries:</b>", bot_map_Ma$data_valid_start,
                           "-", bot_map_Ma$data_valid_end, "<br>",
                           "<b>Chronology:</b>", bot_map_Ma$Chronology, "<br>",
                           "<b>Reference:</b><i>", bot_map_Ma$short_ref, "</i>"
                           ), 
             label=~htmlEscape(bot_map_Ma$site_name)
 ) %>%
  addLayersControl(
    overlayGroups = c("R", "LR", "EMA", "Ma"),
    baseGroups = c("Political", "Physical"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  hideGroup(c("LR", "EMA", "Ma")) %>%
  addResetMapButton() %>%
  addFullscreenControl() %>%
  htmlwidgets::onRender("
    function(el, x) {
      var myMap = this;
      myMap.on('baselayerchange',
        function (e) {
          myMap.minimap.changeLayer(L.tileLayer.provider(e.name));
        })
    }")

```

```{r}
#| echo: false
#| output: false
#| eval: false

# If I want a general map

map_pal2 <- leaflet::colorFactor(palette = c("R" = "#283618", 
                                           "LR" = "#3e9b1c", 
                                           "EMA" = "#bc4749",
                                           "Ma" = "goldenrod"
                                           ), 
                               domain = bot_map$Chronology)

italy_bot_map2 <- leaflet(data=italy) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(weight = 1, 
              smoothFactor = 0.5,
              opacity = 0.2, 
              fillOpacity = 0.5,
              fillColor = "#F9F5EB",
              color="#D7A86E"
              ) %>%
  addCircleMarkers(
    lng=bot_map$x, lat=bot_map$y,
    label = ~htmlEscape(bot_map$site_name),
    color=map_pal2(bot_map$Chronology),
    fillOpacity = 0.5,
    stroke = F,
    radius=4,
     popup = paste("ID:", bot_map$site_code, "<br>",
                           "<b>Site:</b>", bot_map$site_name, "<br>",
                           "Type:", bot_map$type_name, "<br>",
                           "Chronology (c.):", bot_map$startcentury,
                           "-", bot_map$endcentury, "<br>",
                           "Data:", bot_map$available_data, "<br>",
                           "ID BOT:", bot_map$bot, "<br>",
                           "ID ZOO:", bot_map$zoo, "<br>",
                           "ID POLL:", bot_map$poll
                           )
  ) %>% 
  addLegend(position = "bottomright",
            values = bot_map$Chronology, # data frame column for legend
            opacity = .7, # alpha of the legend
            pal = map_pal2, # palette declared earlier
            title = "Data") %>%
  addResetMapButton()
```

::: {.content-visible when-format="html"}
```{r}
#| echo: false
#| label: fig-archaeobot-map
#| fig-cap: "**Legend**: `R` = Roman, `LR` = Late Roman, `EMA` = Early Middle Ages, `Ma` = 11th c. onwards "

italy_bot_map

```
:::

## Ubiquity {style="text-align:justify;"}

::: {.content-visible when-format="html"}
In @sec-methods, ubiquity has been described as the best solution to present the archaeobotanical remains from the Italian peninsula, given the numerous biases in the samples. The heatmap below (@fig-heatmap-ubiquity) provides a good overview of the temporal trends of presence of cereals, legumes, fruits and nuts in the entire area under examination.
:::

::: {.content-visible when-format="pdf"}
In @sec-methods, ubiquity has been described as the best solution to present the archaeobotanical remains from the Italian peninsula, given the numerous biases in the samples. The heatmap below (@fig-heatmap-ubiquity-for-PDF) provides a good overview of the temporal trends of presence of cereals, legumes, fruits and nuts in the entire area under examination.
:::

```{r}
#| echo: false

# SQL file name: view_archaeobot without the Chronology column
#plants_export <- read.csv("/Users/robertoragno/Desktop/Learn #R/PhdTests/PhdTests/DATA/Archaeobotany/PlantsExport14July.csv", header=TRUE, sep=";")
plants_export <- bot_map[-5]

# SQL file name: Archaeobot_Condensed
# Import the file: the function Ubiquity_macroreg_chrono() requires the condensed plant table exported from the database. For more info, look at the Custom functions section.
Archaeobot_Condensed <- read.csv("/Users/robertoragno/Desktop/University/Bari/PhD - Quarto/Database export/Archaeobot_Condensed.csv", header=TRUE, sep=";")

# SQL file name: Archaeobot_viz
# It is the condensed table from the database, but without totals/plant type
Df_Cond_Plants <- read.csv("/Users/robertoragno/Desktop/University/Bari/PhD - Quarto/Database export/Archaeobot_Viz.csv", header=TRUE, sep=";")

```

::: {.content-visible when-format="html"}
```{r}
#| code-fold: true
#| output: false
#| code-summary: "Show the code"
#| code-overflow: wrap

# Load the libraries
# Note: these libraries are used for the data visualizations in this page.
library(RColorBrewer)
library(reshape2)
library(ggplot2)
library(hrbrthemes)
library(plotly)
library(patchwork)

## UBIQUITY

## Creating a dataframe that contains the ubiquity of each century under examination. 
Ubiquity_table <- data.frame(
  "I BCE" = archaeobotany_tables(plants_export, -1)$Ubiquity_exp,  
  "I CE" = archaeobotany_tables(plants_export, 1)$Ubiquity_exp,
  "II CE" = archaeobotany_tables(plants_export, 2)$Ubiquity_exp,
  "III CE" = archaeobotany_tables(plants_export, 3)$Ubiquity_exp,
  "IV CE" = archaeobotany_tables(plants_export, 4)$Ubiquity_exp,
  "V CE" = archaeobotany_tables(plants_export, 5)$Ubiquity_exp,
  "VI CE" = archaeobotany_tables(plants_export, 6)$Ubiquity_exp,
  "VII CE" = archaeobotany_tables(plants_export, 7)$Ubiquity_exp,
  "VIII CE" = archaeobotany_tables(plants_export, 8)$Ubiquity_exp,
  "IX CE" = archaeobotany_tables(plants_export, 9)$Ubiquity_exp,
  "X CE" = archaeobotany_tables(plants_export, 10)$Ubiquity_exp,
  "XI CE" = archaeobotany_tables(plants_export, 11)$Ubiquity_exp
  )

# Transform the ubiquity table into a matrix
Ubiquity_mat <- as.matrix(Ubiquity_table) 

# Rename the centuries
colnames(Ubiquity_mat) <- c("1st c. BCE", "1st c. CE", "2nd c. CE",
                            "3rd c. CE", "4th c. CE", "5th c. CE",
                            "6th c. CE", "7th c. CE", "8th c. CE",
                            "9th c. CE", "10th c. CE", "11th c. CE") 

# The data has to be molten to use it with ggplot2
# (package: reshape2)
Ubiquity_melt <- melt(Ubiquity_mat)

# Let's now rename the columns 
colnames(Ubiquity_melt) <- c("Taxon", "Century", "Ubiquity")

# Add a column for the text tooltip
Ubiquity_melt <- Ubiquity_melt %>%
  mutate(text = paste0("Taxon: ", Taxon, "\n", "Century: ", Century, "\n", "Value: ",round(Ubiquity,2)))

# Create the heatmap with ggplot2
Ubiquity_HM <- ggplot(Ubiquity_melt, aes(Century, Taxon, fill=Ubiquity, text=text)) + 
  geom_tile(colour="white") +
  scale_alpha(range=c(0,1)) +
  scale_x_discrete("", expand = c(0, 0)) + 
  scale_y_discrete("", expand = c(0, 0)) + 
  theme_grey(base_size = 9) + 
  theme(legend.position = "right",
        axis.ticks = element_blank(), 
        axis.text.x = element_text(angle = 90, hjust = 0)
        ) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(
    title="Ubiquity",
    subtitle="Diachronical heatmap of recorded plant species"
  ) +
  scale_fill_gradient(low = "white", high = "black")

```

```{r}
#| echo: false
#| label: fig-heatmap-ubiquity
#| fig-cap: "Diachronical heatmap of recorded plant species" 
#| fig-width: 5
#| fig-height: 6

ggplotly(Ubiquity_HM, tooltip="text")
```
:::

::: {.content-visible when-format="pdf"}
```{r}
#| echo: false
#| label: fig-heatmap-ubiquity-for-PDF
#| fig-cap: "Diachronical heatmap of recorded plant species" 
#| fig-width: 5
#| fig-height: 6

Ubiquity_HM
```
:::

### Ubiquity by macroregion {#sec-macroregional-differences-ubiquity}

::: alert
Probably delete, in favor of the bayesian predictions. If this section is kept, make sure to explain that ubiquity was calculated by macroregion only because it was more reliable etc etc
:::

::: {.content-visible when-format="html"}
The heatmap (@fig-heatmap-ubiquity) shows the diachronical ubiquity values of the Italian mainland. It is possible however to compare plants' ubiquity values in different areas of the peninsula. The R function `Ubiquity_macroreg_chrono()` (@sec-Ubiquity-macroreg-chrono) was created to subset data from Northern, Central and Southern Italian regions, using modern boundaries. Archaeobotanical data from Italy is scarce, and subsetting the dataset required a larger chronological division to have enough sites for a valid statistical interpretation of the results. For this reason, the ubiquity values are presented using the variable `Chronology` rather than subsetting the individual centuries. For a clearer reading of the plot, the taxa have been divided into--Cereals, Pulses and Fruits/Nuts. Some taxa have been omitted from the plot.
:::

::: {.content-visible when-format="pdf"}
The heatmap (@fig-heatmap-ubiquity-for-PDF) shows the diachronical ubiquity values of the Italian mainland. It is possible however to compare plants' ubiquity values in different areas of the peninsula. The R function `Ubiquity_macroreg_chrono()` (@sec-custom-functions) was created to subset data from Northern, Central and Southern Italian regions, using modern boundaries. Archaeobotanical data from Italy is scarce, and subsetting the dataset required a larger chronological division to have enough sites for a valid statistical interpretation of the results. For this reason, the ubiquity values are presented using the variable `Chronology` rather than subsetting the individual centuries. For a clearer reading of the plot, the taxa have been divided into--Cereals, Pulses and Fruits/Nuts. Some taxa have been omitted from the plot.
:::

::: {.content-visible when-format="html"}
```{r}
#| echo: false
#| output: false
#| code-summary: "Show the code: data preparation"
#| code-overflow: wrap

# Ubiquity by Italian Macro regions: Northern, Central and Southern Italy

# Load the libraries
library(vegan)
library(matrixStats)
library(patchwork)

# Creating a dataframe with the ubiquities of all macroregions and chronologies
bot_macroreg <- rbind(
  Ubiquity_R_NI <- Ubiquity_macroreg_chrono(Archaeobot_Condensed,"Northern Italy", "R"),
  Ubiquity_R_CI <- Ubiquity_macroreg_chrono(Archaeobot_Condensed,"Central Italy", "R"),
  Ubiquity_R_SI <- Ubiquity_macroreg_chrono(Archaeobot_Condensed,"Southern Italy", "R"),
  Ubiquity_LR_NI <- Ubiquity_macroreg_chrono(Archaeobot_Condensed,"Northern Italy", "LR"),
  Ubiquity_LR_CI <- Ubiquity_macroreg_chrono(Archaeobot_Condensed,"Central Italy", "LR"),
  Ubiquity_LR_SI <- Ubiquity_macroreg_chrono(Archaeobot_Condensed,"Southern Italy", "LR"),
  Ubiquity_EMA_NI <- Ubiquity_macroreg_chrono(Archaeobot_Condensed,"Northern Italy", "EMA"),
  Ubiquity_EMA_CI <- Ubiquity_macroreg_chrono(Archaeobot_Condensed,"Central Italy", "EMA"),
  Ubiquity_EMA_SI <- Ubiquity_macroreg_chrono(Archaeobot_Condensed,"Southern Italy", "EMA"),
  Ubiquity_Ma_NI <- Ubiquity_macroreg_chrono(Archaeobot_Condensed,"Northern Italy", "Ma"),
  Ubiquity_Ma_CI <- Ubiquity_macroreg_chrono(Archaeobot_Condensed,"Central Italy", "Ma"),
  Ubiquity_Ma_SI <- Ubiquity_macroreg_chrono(Archaeobot_Condensed,"Southern Italy", "Ma")
)

# Re-arranging the cereals/macroregions for visualisation on the Y axis
level_macroreg_order <- c("Southern Italy", "Central Italy", "Northern Italy")

level_cereals_order <- c("Common.Wheat", "Barley", "Rye", 
                         "Einkorn", "Emmer", "Proso.millet", 
                         "Foxtail.millet", "Oats", "Sorghum")

# Cereals
cer_ubiquity_macroreg.R <- filter(bot_macroreg, Chronology=="R" & Plant.Type=="Cereals")
cer_ubiquity_macroreg.R <- filter(cer_ubiquity_macroreg.R, Macroregion!="Central Italy")
cer_ubiquity_macroreg.LR <- filter(bot_macroreg, Chronology=="LR" & Plant.Type=="Cereals")
cer_ubiquity_macroreg.EMA <- filter(bot_macroreg, Chronology=="EMA" & Plant.Type=="Cereals")
cer_ubiquity_macroreg.Ma <- filter(bot_macroreg, (Chronology=="Ma" & Plant.Type=="Cereals"))
cer_ubiquity_macroreg.Ma <- filter(cer_ubiquity_macroreg.Ma, Macroregion!="Southern Italy")

#Pulses
puls_ubiquity_macroreg.R <- filter(bot_macroreg, Chronology=="R" & Plant.Type=="Pulses")
puls_ubiquity_macroreg.R <- filter(puls_ubiquity_macroreg.R, Macroregion!="Central Italy")
puls_ubiquity_macroreg.R <- filter(puls_ubiquity_macroreg.R, Plant!="Chickpea")
puls_ubiquity_macroreg.LR <- filter(bot_macroreg, Chronology=="LR" & Plant.Type=="Pulses")
puls_ubiquity_macroreg.LR <- filter(puls_ubiquity_macroreg.LR, 
                                    Macroregion!="Southern Italy")
puls_ubiquity_macroreg.LR <- filter(puls_ubiquity_macroreg.LR, Plant!="Chickpea")
puls_ubiquity_macroreg.EMA <- filter(bot_macroreg, Chronology=="EMA" & Plant.Type=="Pulses")
puls_ubiquity_macroreg.Ma <- filter(bot_macroreg, Chronology=="Ma"  & Plant.Type=="Pulses")
puls_ubiquity_macroreg.Ma <- filter(puls_ubiquity_macroreg.Ma, 
                                    Macroregion!="Southern Italy")

#Fruits (+ Subset)
fnuts_ubiquity_macroreg.R <- filter(bot_macroreg, Chronology=="R" & Plant.Type=="Fruits/Nuts")
fnuts_ubiquity_macroreg.R <- subset(fnuts_ubiquity_macroreg.R, (Plant == "Wild.Cherry" | Plant == "Walnut" | Plant == "Peach" | Plant == "Olive" |Plant == "Grape" | Plant =="Fig" | Plant =="Apple"))
fnuts_ubiquity_macroreg.R <- filter(fnuts_ubiquity_macroreg.R, Macroregion!="Central Italy")
fnuts_ubiquity_macroreg.LR <- filter(bot_macroreg, Chronology=="LR" & Plant.Type=="Fruits/Nuts")
fnuts_ubiquity_macroreg.LR <- subset(fnuts_ubiquity_macroreg.LR, (Plant == "Wild.Cherry" | Plant == "Walnut" | Plant == "Peach" | Plant == "Olive" |Plant == "Grape" | Plant =="Fig" | Plant =="Apple"))
fnuts_ubiquity_macroreg.EMA <- filter(bot_macroreg, Chronology=="EMA" & Plant.Type=="Fruits/Nuts")
fnuts_ubiquity_macroreg.EMA <- subset(fnuts_ubiquity_macroreg.EMA, (Plant == "Wild.Cherry" | Plant == "Walnut" | Plant == "Peach" | Plant == "Olive" |Plant == "Grape" | Plant =="Fig" | Plant =="Apple"))
fnuts_ubiquity_macroreg.Ma <- filter(bot_macroreg, Chronology=="Ma"  & Plant.Type=="Fruits/Nuts")
fnuts_ubiquity_macroreg.Ma <- subset(fnuts_ubiquity_macroreg.Ma, (Plant == "Wild.Cherry" | Plant == "Walnut" | Plant == "Peach" | Plant == "Olive" |Plant == "Grape" | Plant =="Fig" | Plant =="Apple"))
fnuts_ubiquity_macroreg.Ma <- filter(fnuts_ubiquity_macroreg.Ma, Macroregion!="Southern Italy")

```

```{r}
#| echo: false
#| output: false
#| code-summary: "Show the code: plots"
#| code-overflow: wrap

# Cereals plots ubiquity
cer_ubiquity_macroreg_R.HM <- ggplot(cer_ubiquity_macroreg.R, aes(
  factor(Macroregion, levels=(level_macroreg_order)),
  factor(Plant, levels=rev(level_cereals_order)),
  fill=(Ubiquity)
)) + 
  geom_tile(colour="white") +
  geom_text(aes(label = Ubiquity), colour="white", size=3)+ 
  scale_alpha(range=c(0,1)) +
  scale_x_discrete("", expand = c(0, 0)) + 
  scale_y_discrete("", expand = c(0, 0)) + 
  theme_grey(base_size = 9) + 
  theme(legend.position = "none",
        axis.ticks = element_blank()
  ) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(
    title="Roman"
  ) +  scale_fill_gradient(low = "white", high = "black")


cer_ubiquity_macroreg_LR.HM <- ggplot(cer_ubiquity_macroreg.LR, aes(
  factor(Macroregion, levels=(level_macroreg_order)),
  factor(Plant, levels=rev(level_cereals_order)),
  fill=(Ubiquity)
)) + 
  geom_tile(colour="white") +
  geom_text(aes(label = Ubiquity), colour="white", size=3)+ 
  scale_alpha(range=c(0,1)) +
  scale_x_discrete("", expand = c(0, 0)) + 
  scale_y_discrete("", expand = c(0, 0)) + 
  theme_grey(base_size = 9) + 
  theme(legend.position = "none",
        axis.ticks = element_blank()
  ) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(
    title="Late Roman"
  ) +  scale_fill_gradient(low = "white", high = "black")

cer_ubiquity_macroreg_EMA.HM <- ggplot(cer_ubiquity_macroreg.EMA, aes(
  factor(Macroregion, levels=(level_macroreg_order)),
  factor(Plant, levels=rev(level_cereals_order)),
  fill=(Ubiquity)
)) + 
  geom_tile(colour="white") +
  geom_text(aes(label = Ubiquity), colour="white", size=3)+ 
  scale_alpha(range=c(0,1)) +
  scale_x_discrete("", expand = c(0, 0)) + 
  scale_y_discrete("", expand = c(0, 0)) + 
  theme_grey(base_size = 9) + 
  theme(legend.position = "none",
        axis.ticks = element_blank()
  ) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(
    title="Early Medieval"
  ) +  scale_fill_gradient(low = "white", high = "black")

cer_ubiquity_macroreg_Ma.HM <- ggplot(cer_ubiquity_macroreg.Ma, aes(
  factor(Macroregion, levels=(level_macroreg_order)),
  factor(Plant, levels=rev(level_cereals_order)),
  fill=(Ubiquity)
)) + 
  geom_tile(colour="white") +
  geom_text(aes(label = Ubiquity), colour="white", size=3)+ 
  scale_alpha(range=c(0,1)) +
  scale_x_discrete("", expand = c(0, 0)) + 
  scale_y_discrete("", expand = c(0, 0)) + 
  theme_grey(base_size = 9) + 
  theme(legend.position = "none",
        axis.ticks = element_blank()
  ) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(
    title="Medieval"
  ) +  scale_fill_gradient(low = "white", high = "black")


Cereals_Ubiquity_MacroReg_Patchwork <- (cer_ubiquity_macroreg_R.HM|cer_ubiquity_macroreg_LR.HM)/(cer_ubiquity_macroreg_EMA.HM|cer_ubiquity_macroreg_Ma.HM)
Cereals_Ubiquity_MacroReg_Patchwork + plot_annotation(
  title = 'Cereals',
  subtitle = 'Ubiquity (%), plotted by macroregion and chronology.',
  caption='Note: Data was too scarce for Roman Central Italy and Medieval Southern Italy.'
  )


# Pulses plots ubiquity
puls_ubiquity_macroreg_R.HM <- ggplot(puls_ubiquity_macroreg.R, aes(
  factor(Macroregion, levels=(level_macroreg_order)),
  Plant,
  fill=(Ubiquity)
)) + 
  geom_tile(colour="white") +
  geom_text(aes(label = Ubiquity), colour="#cfcfcf", size=3)+ 
  scale_alpha(range=c(0,1)) +
  scale_x_discrete("", expand = c(0, 0)) + 
  scale_y_discrete("", expand = c(0, 0)) + 
  theme_grey(base_size = 9) + 
  theme(legend.position = "none",
        axis.ticks = element_blank()
  ) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(
    title="Roman"
  ) +  scale_fill_gradient(low = "white", high = "black")


puls_ubiquity_macroreg_LR.HM <- ggplot(puls_ubiquity_macroreg.LR, aes(
  factor(Macroregion, levels=(level_macroreg_order)),
  Plant,
  fill=(Ubiquity)
)) + 
  geom_tile(colour="white") +
  geom_text(aes(label = Ubiquity), colour="#ffffff", size=3)+ 
  scale_alpha(range=c(0,1)) +
  scale_x_discrete("", expand = c(0, 0)) + 
  scale_y_discrete("", expand = c(0, 0)) + 
  theme_grey(base_size = 9) + 
  theme(legend.position = "none",
        axis.ticks = element_blank()
  ) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(
    title="Late Roman"
  ) +  scale_fill_gradient(low = "white", high = "black")

puls_ubiquity_macroreg_EMA.HM <- ggplot(puls_ubiquity_macroreg.EMA, aes(
  factor(Macroregion, levels=(level_macroreg_order)),
  Plant,
  fill=(Ubiquity)
)) + 
  geom_tile(colour="white") +
  geom_text(aes(label = Ubiquity), colour="white", size=3)+ 
  scale_alpha(range=c(0,1)) +
  scale_x_discrete("", expand = c(0, 0)) + 
  scale_y_discrete("", expand = c(0, 0)) + 
  theme_grey(base_size = 9) + 
  theme(legend.position = "none",
        axis.ticks = element_blank()
  ) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(
    title="Early Medieval"
  ) +  scale_fill_gradient(low = "white", high = "black")

puls_ubiquity_macroreg_Ma.HM <- ggplot(puls_ubiquity_macroreg.Ma, aes(
  factor(Macroregion, levels=(level_macroreg_order)),
  Plant,
  fill=(Ubiquity)
)) + 
  geom_tile(colour="white") +
  geom_text(aes(label = Ubiquity), colour="white", size=3)+ 
  scale_alpha(range=c(0,1)) +
  scale_x_discrete("", expand = c(0, 0)) + 
  scale_y_discrete("", expand = c(0, 0)) + 
  theme_grey(base_size = 9) + 
  theme(legend.position = "none",
        axis.ticks = element_blank()
  ) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(
    title="Medieval"
  ) +  scale_fill_gradient(low = "white", high = "black")


Pulses_Ubiquity_MacroReg_Patchwork <- (puls_ubiquity_macroreg_R.HM|puls_ubiquity_macroreg_LR.HM)/(puls_ubiquity_macroreg_EMA.HM|puls_ubiquity_macroreg_Ma.HM)
Pulses_Ubiquity_MacroReg_Patchwork + plot_annotation(
  title = 'Pulses',
  subtitle = 'Ubiquity (%), plotted by macroregion and chronology.',
  caption='Note: Data was too scarce for Roman Central Italy and Late Roman/Medieval Southern Italy.'
)

# Fruits nuts plots

fnuts_ubiquity_macroreg_R.HM <- ggplot(fnuts_ubiquity_macroreg.R, aes(
  factor(Macroregion, levels=(level_macroreg_order)),
  Plant,
  fill=(Ubiquity)
)) + 
  geom_tile(colour="white") +
  geom_text(aes(label = Ubiquity), colour="white", size=3)+ 
  scale_alpha(range=c(0,1)) +
  scale_x_discrete("", expand = c(0, 0)) + 
  scale_y_discrete("", expand = c(0, 0)) + 
  theme_grey(base_size = 9) + 
  theme(legend.position = "none",
        axis.ticks = element_blank()
  ) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(
    title="Roman"
  ) +  scale_fill_gradient(low = "white", high = "black")


fnuts_ubiquity_macroreg_LR.HM <- ggplot(fnuts_ubiquity_macroreg.LR, aes(
  factor(Macroregion, levels=(level_macroreg_order)),
  Plant,
  fill=(Ubiquity)
)) + 
  geom_tile(colour="white") +
  geom_text(aes(label = Ubiquity), colour="white", size=3)+ 
  scale_alpha(range=c(0,1)) +
  scale_x_discrete("", expand = c(0, 0)) + 
  scale_y_discrete("", expand = c(0, 0)) + 
  theme_grey(base_size = 9) + 
  theme(legend.position = "none",
        axis.ticks = element_blank()
  ) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(
    title="Late Roman"
  ) +  scale_fill_gradient(low = "white", high = "black")

fnuts_ubiquity_macroreg_EMA.HM <- ggplot(fnuts_ubiquity_macroreg.EMA, aes(
  factor(Macroregion, levels=(level_macroreg_order)),
  Plant,
  fill=(Ubiquity)
)) + 
  geom_tile(colour="white") +
  geom_text(aes(label = Ubiquity), colour="white", size=3)+ 
  scale_alpha(range=c(0,1)) +
  scale_x_discrete("", expand = c(0, 0)) + 
  scale_y_discrete("", expand = c(0, 0)) + 
  theme_grey(base_size = 9) + 
  theme(legend.position = "none",
        axis.ticks = element_blank()
  ) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(
    title="Early Medieval"
  ) +  scale_fill_gradient(low = "white", high = "black")

fnuts_ubiquity_macroreg_Ma.HM <- ggplot(fnuts_ubiquity_macroreg.Ma, aes(
  factor(Macroregion, levels=(level_macroreg_order)),
  Plant,
  fill=(Ubiquity)
)) + 
  geom_tile(colour="white") +
  geom_text(aes(label = Ubiquity), colour="white", size=3)+ 
  scale_alpha(range=c(0,1)) +
  scale_x_discrete("", expand = c(0, 0)) + 
  scale_y_discrete("", expand = c(0, 0)) + 
  theme_grey(base_size = 9) + 
  theme(legend.position = "none",
        axis.ticks = element_blank()
  ) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  labs(
    title="Medieval"
  ) +  scale_fill_gradient(low = "white", high = "black")


FrNuts_Ubiquity_MacroReg_Patchwork <- (fnuts_ubiquity_macroreg_R.HM|fnuts_ubiquity_macroreg_LR.HM)/(fnuts_ubiquity_macroreg_EMA.HM|fnuts_ubiquity_macroreg_Ma.HM)
FrNuts_Ubiquity_MacroReg_Patchwork + plot_annotation(
  title = 'Fruits/Nuts',
  subtitle = 'Ubiquity (%), plotted by macroregion and chronology.',
  caption='Note: Data was too scarce for Roman Central Italy and Medieval Southern Italy.'
)


```
:::

#### Cereals

It is interesting to notice how in the **Roman age**, cereals are similarly ubiquitous in Southern and Northern Italy, although there are some exceptions (*i.e.* einkorn, rye, oats, and proso millet) that can derive from the randomness of sampling. Unfortunately, only three sites provided botanical samples for Roman Central Italy and their values have then been omitted from the plot. These sites (all from Tuscany) were studied by the Roman Peasant Project [@bowesRomanPeasantProject2020a] and only reported three kinds of cereal: common wheat, emmer, and barley. Similar ubiquity values from Northern and Southern Italy in the Roman age may suggest similar production patterns in the whole Italian mainland, even though more data is required. In the **Late Roman age**, ubiquity data has been calculated for the three macroregions, with Southern Italy being the least trustworthy (five sites in total). Three crops are found on 62.5-75% of the Central Italian sites: common wheat, barley and emmer. Other cereals are present, but less ubiquitously. The cereal triad aforementioned seems to be diffused in the south as well. Conversely, in Northern Italy common wheat and barley were indeed important cultivations but had to compete with other cereals including millet, sorghum, and rye. The latter had a significant increase, being present on almost 30% of the Northern sites (as opposed to the Roman 15%). The **Early Medieval age** seems to mark a shift in Italian agricultural practices, as cereal ubiquities are much more variable regionally. In Southern Italy, the triad of common wheat, barley and emmer were still the predominant cereals. These cereals are ubiquitous in Central and Northern Italy as well, although these regions adopt polyculture with a diversified number of cereals. The samples from the **Medieval age** are fewer in number since the upper boundary of this project's chronology is the 11^th^ c. Despite the short chronology, it is possible to make some considerations. Medieval Centraly Italy relied heavily on common wheat, barley and emmer, with other cereals increasingly important. Barley is the most ubiquitous cereal in Northern Italy in this period, followed by common wheat, millets and sorghum.

```{r}
#| echo: false
#| label: fig-heatmap-ubiquity-cereals-macroreg
#| fig-cap: "Diachronical heatmap of cereals in the Italian macroregions"
#| fig-dpi: 300 

Cereals_Ubiquity_MacroReg_Patchwork

```

#### Pulses

In the **Roman Age**, pulses are an important part of the diet and are cultivated both in Northern and Southern Italy. In the latter, vetch/broad beans are present in 22-32% of the samples, and lentils are present in 38% of the sites. In the **Late Roman Age**, broad beans are equally important in Central and Northern Italy, and peas are present in 50% of the Central Italian sites. In the **Early Medieval Age**, pulses are present in many Central Italian sites, especially blue/red peas, broad beans and other Fabaceae. Lentils and broad beans are also cultivated in almost half of the Northern Italian sites. The importance of pulses in Central Italy is confirmed by the **11^th^ c.** samples, where every specie is present in over 66% of the sites and Fabaceae and blue/red peas are found in every sample. Conversely, in Northern Italy broad bean is found in 66% of the sites.

```{r}
#| echo: false
#| label: fig-heatmap-ubiquity-pulses-macroreg
#| fig-cap: "Diachronical heatmap of pulses in the Italian macroregions" 

Pulses_Ubiquity_MacroReg_Patchwork

```

#### Fruits and nuts

Olive and grape are two essential cultivations in the Italian peninsula. Olive pits, as can be expected, are more ubiquitous in Southern Italy, where in Roman times are present in \>87% of the sites and in over 58% of the sites in the following chronologies[^archaeobotany-1]. Conversely, grape is important in Central and Northern Italy in the Late Roman, Early Medieval and Medieval ages.

[^archaeobotany-1]: The Late Roman values for Southern Italy are only based on 5 samples (3 of which are from the same site, Salapia) so the values are not very trustworthy.

```{r}
#| echo: false
#| label: fig-heatmap-ubiquity-fruit-macroreg
#| fig-cap: "Diachronical heatmap of fruits/nuts in the Italian macroregions" 

FrNuts_Ubiquity_MacroReg_Patchwork

```

# Context type {style="text-align:justify"}

::: callout-note
Important: This section has to be rewritten as now the approach is not frequentist anymore.
:::

It is possible to examine the distribution of plants across different site types during the four phases under examination. During the Roman period, wheat and barley were most prevalent on rural and religious sites, in the latter being used as part of ritual offerings. Minor grains were widely distributed on religious and urban sites, with rural sites and villas also showing high percentages. Legumes were also used in religious offerings but were found on both urban and rural sites. Grapes were ubiquitous across all Roman sites, with particularly high percentages in urban contexts and being extensively cultivated in large Roman estates. Olives had lower percentages, as they cannot be grown in Northern Italy, where most of the samples were collected. However, they were still widely distributed across all contexts, with peak percentages in urban and religious sites. In the Late Roman period, olive presence significantly decreased in urban sites and villas, but remained relatively unchanged in rural contexts. Similarly to olives, grape decrease in the same contexts, but slightly increase in rural sites. In addition to rural sites, grapes are also diffused in over 66% of the fortified sites (*castra*) that start to appear in this period. It is noteworthy that fortified sites from the Late Roman period exhibit high percentages of both minor and noble grains, possibly being stocked with the addition of pulses such as faba beans and vetch. In the EMA phase, noble/minor grains and pulses are still present in almost any fortified sites, a situation that is similar to the one in rural sites. In rural sites however minor grains (40%) and pulses (32%) are less ubiquitous in this period, although this figure might suffer from sites that only published cereal remains. Grapes are present in every religious/funerary context, followed by fortified (71%) and rural sites (66%). If olive cultivation seem to decrease in this phase, it might be because most of the sites with archaeobotanical remains are located in Northern Italy. Olives are present in 50% of the religious sites and on 36% of villa sites. Olives are absent from fortified sites in all phases, probably as a reflection of the northern location of these sites. Both rural and urban sites increase the consumption of berries in this period (13-14% ca.), which are also used for funerary offerings (14%).

```{r}
#| echo: false
#| eval: false
                                                                
# Store a copy of the original dataframe
Plants_Df_for_Ubiquity <- Df_Cond_Plants 

# Simplify categories
#table(Plants_Df_for_Ubiquity$Type)
Plants_Df_for_Ubiquity$Type <- str_replace(Plants_Df_for_Ubiquity$Type, "Religious, monastery", "Religious")    
Plants_Df_for_Ubiquity$Type <- str_replace(Plants_Df_for_Ubiquity$Type, "Castle", "Fortified")    
Plants_Df_for_Ubiquity$Type <- str_replace(Plants_Df_for_Ubiquity$Type, "Castrum", "Fortified")  

# Remove unsp.. columns
Plants_Df_for_Ubiquity <- Plants_Df_for_Ubiquity %>% select(-"Unsp..cereals", -"Unsp..Pulses")

# Define types and chrono
Roman_Site_types <- c("Necropolis", "Religious", "Rural", "Rural site, villa", "Urban")
LateRoman_Site_types <- c("Fortified", "Necropolis", "Rural", "Rural site, villa", "Urban")
EMA_Site_types <- c("Fortified", "Necropolis", "Religious", "Rural", "Rural site, villa", "Urban")
Ma_Site_types <- c("Fortified", "Religious", "Rural", "Urban")

# Define a list of column groups
plants_groups <- list(
  Noble.Grains = c("Common.Wheat", "Barley"),
  Minor.Grains = c("Emmer", "Einkorn", "Oats", "Rye", "Proso.millet", "Foxtail.millet", "Sorghum"),
  Legumes= c("Lentil","Pea","Faba.bean","Vetch","Unsp..Vicia","Blue.Red.Pea","Chickpea"),
  Nuts = c("Walnut", "Hazelnut", "Chestnut"),
  Pome.fruits = c("Apple", "Pear"),
  Stone.fruits = c("Peach", "Plum", "Sour.Cherry", "Unsp..Prunus"),
  Berries = c("Blackberry", "Elderberry", "Blackthorn", "Wild.Cherry", "Corn..Cherry", "Sorbus.sp.", "Strawberry"),
  Other.fruits = c("Date", "Melon", "Fig")
)

# Chronology: Roman (R)
Ubiquity_typology_chronology.Roman <- Roman_Site_types %>% 
  map_df(~create_ubiquity_df(Plants_Df_for_Ubiquity, .x, "R"))

# Iterate over the column groups
for (name in names(plants_groups)) {
  
  # Calculate the mean of the columns in the group
  Ubiquity_typology_chronology.Roman <- Ubiquity_typology_chronology.Roman %>%
    mutate(!!name := rowMeans(select(., !!plants_groups[[name]]))) %>% 
                      select(-!!plants_groups[[name]])
}

Ubiquity_typology_chronology.Roman$Chronology <- "R"
Ubiquity_typology_chronology.Roman$Type <- rownames(Ubiquity_typology_chronology.Roman)
rownames(Ubiquity_typology_chronology.Roman) <- NULL

# Chronology: Late Roman (LR)
Ubiquity_typology_chronology.LateRoman <- LateRoman_Site_types %>% 
  map_df(~create_ubiquity_df(Plants_Df_for_Ubiquity, .x, "LR"))

# Iterate over the column groups
for (name in names(plants_groups)) {
  
  # Calculate the mean of the columns in the group
  Ubiquity_typology_chronology.LateRoman <- Ubiquity_typology_chronology.LateRoman %>%
    mutate(!!name := rowMeans(select(., !!plants_groups[[name]]))) %>% 
                      select(-!!plants_groups[[name]])
}

Ubiquity_typology_chronology.LateRoman$Chronology <- "LR"
Ubiquity_typology_chronology.LateRoman$Type <- rownames(Ubiquity_typology_chronology.LateRoman)
rownames(Ubiquity_typology_chronology.LateRoman) <- NULL

# Chronology: Early Medieval (EMA)
Ubiquity_typology_chronology.EMA <- EMA_Site_types %>% 
  map_df(~create_ubiquity_df(Plants_Df_for_Ubiquity, .x, "EMA"))

# Iterate over the column groups
for (name in names(plants_groups)) {
  
  # Calculate the mean of the columns in the group
  Ubiquity_typology_chronology.EMA <- Ubiquity_typology_chronology.EMA %>%
    mutate(!!name := rowMeans(select(., !!plants_groups[[name]]))) %>% 
                      select(-!!plants_groups[[name]])
}

Ubiquity_typology_chronology.EMA$Chronology <- "EMA"
Ubiquity_typology_chronology.EMA$Type <- rownames(Ubiquity_typology_chronology.EMA)
rownames(Ubiquity_typology_chronology.EMA) <- NULL

# Chronology: Medieval (Ma)
Ubiquity_typology_chronology.Ma <- Ma_Site_types %>% 
  map_df(~create_ubiquity_df(Plants_Df_for_Ubiquity, .x, "Ma"))

# Iterate over the column groups
for (name in names(plants_groups)) {
  
  # Calculate the mean of the columns in the group
  Ubiquity_typology_chronology.Ma <- Ubiquity_typology_chronology.Ma %>%
    mutate(!!name := rowMeans(select(., !!plants_groups[[name]]))) %>% 
                      select(-!!plants_groups[[name]])
}

Ubiquity_typology_chronology.Ma$Chronology <- "Ma"
Ubiquity_typology_chronology.Ma$Type <- rownames(Ubiquity_typology_chronology.Ma)
rownames(Ubiquity_typology_chronology.Ma) <- NULL


Ubiquity_All_Chrono_Type <- rbind(
  Ubiquity_typology_chronology.Roman,
  Ubiquity_typology_chronology.LateRoman,
  Ubiquity_typology_chronology.EMA,
  Ubiquity_typology_chronology.Ma
)

# Reorder columns
Ubiquity_All_Chrono_Type <- Ubiquity_All_Chrono_Type %>%
  select("Chronology", "Type", "Noble.Grains","Minor.Grains","Legumes","Nuts","Pome.fruits","Stone.fruits","Berries","Other.fruits","Flax","Grape","Olive")

# PLOTS
level_order <- c("R", "LR", "EMA", "Ma") 
library(ggpubr)

noblegrains_by_type <- ggplot(data=Ubiquity_All_Chrono_Type, 
       aes(
         fill=factor(Chronology, levels=level_order),
         x=Type,
         y=Noble.Grains, 
         ), 
       ) +
        geom_bar(position="dodge", stat="identity") + 
        theme_pubclean() + 
        theme(legend.position="top") +
        scale_fill_grey() +
        labs(
          title = "Noble Grains",
          x= "",
          y = "%",
        fill="Chronology"
    )

minorgrains_by_type <- ggplot(data=Ubiquity_All_Chrono_Type, 
       aes(
         fill=factor(Chronology, levels=level_order),
         x=Type,
         y=Minor.Grains, 
         ), 
       ) +
        geom_bar(position="dodge", stat="identity") + 
        theme_pubclean() + 
        theme(legend.position="none") +
        scale_fill_grey() +
        labs(
          title = "Minor Grains",
          x= "",
          y = "%",
        fill="Chronology"
    )

legumes_by_type <- ggplot(data=Ubiquity_All_Chrono_Type, 
       aes(
         fill=factor(Chronology, levels=level_order),
         x=Type,
         y=Legumes, 
         ), 
       ) +
        geom_bar(position="dodge", stat="identity") + 
        theme_pubclean() + 
        theme(legend.position="none") +
        scale_fill_grey() +
        labs(
          title = "Legumes",
          x= "",
          y = "%",
        fill="Chronology"
    )

grape_by_type <- ggplot(data=Ubiquity_All_Chrono_Type, 
       aes(
         fill=factor(Chronology, levels=level_order),
         x=Type,
         y=Grape, 
         ), 
       ) +
        geom_bar(position="dodge", stat="identity") + 
        theme_pubclean() + 
        theme(legend.position="none") +
        scale_fill_grey() +
        labs(
          title = "Grape",
          x= "",
          y = "%",
        fill="Chronology"
    )

olive_by_type <- ggplot(data=Ubiquity_All_Chrono_Type, 
       aes(
         fill=factor(Chronology, levels=level_order),
         x=Type,
         y=Olive, 
         ), 
       ) +
        geom_bar(position="dodge", stat="identity") + 
        theme_pubclean() + 
        theme(legend.position="none") +
        scale_fill_grey() +
        labs(
          title = "Olives",
          x= "",
          y = "%",
        fill="Chronology"
    )

nuts_by_type <- ggplot(data=Ubiquity_All_Chrono_Type, 
       aes(
         fill=factor(Chronology, levels=level_order),
         x=Type,
         y=Nuts, 
         ), 
       ) +
        geom_bar(position="dodge", stat="identity") + 
        theme_pubclean() + 
        theme(legend.position="none") +
        scale_fill_grey() +
        labs(
          title = "Nuts",
          x= "",
          y = "%",
        fill="Chronology"
    )

pomefruits_by_type <- ggplot(data=Ubiquity_All_Chrono_Type, 
       aes(
         fill=factor(Chronology, levels=level_order),
         x=Type,
         y=Pome.fruits, 
         ), 
       ) +
        geom_bar(position="dodge", stat="identity") + 
        theme_pubclean() + 
        theme(legend.position="bottom") +
        scale_fill_grey() +
        labs(
          title = "Pome Fruits",
          x= "",
          y = "%",
        fill="Chronology"
    )

stonefruits_by_type <- ggplot(data=Ubiquity_All_Chrono_Type, 
       aes(
         fill=factor(Chronology, levels=level_order),
         x=Type,
         y=Stone.fruits, 
         ), 
       ) +
        geom_bar(position="dodge", stat="identity") + 
        theme_pubclean() + 
        theme(legend.position="none") +
        scale_fill_grey() +
        labs(
          title = "Stone Fruits",
          x= "",
          y = "%",
        fill="Chronology"
    )

```

```{r}
#| echo: false
#| warning: false
#| message: false

# Note: The code has to be cleaned because this is a new update and it is not related to the plots used just above this code. 
# For the moment, I will start from scratch
library(dplyr)

Df_Cond_Plants.pa <- Df_Cond_Plants
Df_Cond_Plants.pa[,c(11:50)] <- decostand(Df_Cond_Plants.pa[,c(11:50)], method="pa")

#Simplify categories
Df_Cond_Plants.pa$Type <- str_replace(Df_Cond_Plants.pa$Type, "Religious, monastery", "Religious")    
Df_Cond_Plants.pa$Type <- str_replace(Df_Cond_Plants.pa$Type, "Castle", "Fortified")    
Df_Cond_Plants.pa$Type <- str_replace(Df_Cond_Plants.pa$Type, "Castrum", "Fortified")  

```

```{r}
#| echo: false
#| eval: true
#| code-fold: true

Minor_Grains <- with(
  Df_Cond_Plants.pa,
  data.frame(
  Chronology = Chronology,
  From.Century=From.Century, 
  To.Century=To.Century,
  Macroregion = name_macroreg,
  Site_Type = as.factor(Type),
  Geo = Geo,
  Grain_Count = Einkorn+Oats+Rye+Proso.millet+Foxtail.millet+Sorghum+Emmer,
  All_Grain_Count = rowSums(Df_Cond_Plants.pa[,c(11:19)])
  )
)

# Transform categories to factors
Minor_Grains[,c(1,4,5,6)] <- lapply(Minor_Grains[,c(1,4,5,6)], factor)

Noble_Grains_Reg <- with(
  Df_Cond_Plants.pa,
  data.frame(
  Chronology = Chronology,
  From.Century=From.Century, 
  To.Century=To.Century,
  Macroregion = name_macroreg,
  Site_Type = as.factor(Type),
  Geo = Geo,
  Grain_Count = Common.Wheat+Barley,
  All_Grain_Count = rowSums(Df_Cond_Plants.pa[,c(11:19)])
  )
)

# Transform categories to factors
Noble_Grains_Reg[,c(1,4,5,6)] <- lapply(Noble_Grains_Reg[,c(1,4,5,6)], factor)


```

```{r}
#| echo: true
#| code-fold: true
#| eval: true
#| message: false
#| output: false

library(rethinking)
library(tidybayes.rethinking)
library(tidybayes)

####################
## NOBLE GRAINS
####################

Noble_Grains <- with(
  Df_Cond_Plants.pa,
  data.frame(
  Chronology = Chronology,
  Macroregion = name_macroreg,
  Site_Type = as.factor(Type),
  Geo = Geo,
  Grain_Count = rowSums(Df_Cond_Plants.pa[,c(11,15)])
  )
)

#Simplify categories
Noble_Grains$Site_Type <- str_replace(Noble_Grains$Site_Type, "Religious, monastery", "Religious")    
Noble_Grains$Site_Type <- str_replace(Noble_Grains$Site_Type, "Castle", "Fortified")    
Noble_Grains$Site_Type <- str_replace(Noble_Grains$Site_Type, "Castrum", "Fortified")  

# Convert to list
Noble_Grain.list <- list(
    NG_Found = Noble_Grains$Grain_Count,
    NG_Tot = 2,
    TC_ID = as.factor((interaction(Noble_Grains$Chronology, 
                                   Noble_Grains$Site_Type)))
   )

Noble_Grain.list$TC_ID <- droplevels(Noble_Grain.list$TC_ID)


####################
## MINOR GRAINS
####################

# Convert to list
Minor_Grains_Type.list <- list(
    MG_Found = Minor_Grains$Grain_Count,
    MG_Tot = 7,
    TC_ID = as.factor((interaction(Minor_Grains$Chronology, 
                                   Minor_Grains$Site_Type)))
   )

Minor_Grains_Type.list$TC_ID <- droplevels(Minor_Grains_Type.list$TC_ID)
  

####################
## LEGUMES 
####################

Legumes <- with(
  Df_Cond_Plants.pa,
  data.frame(
  Chronology = Chronology,
  Macroregion = name_macroreg,
  Site_Type = as.factor(Type),
  Geo = Geo,
  Seed_Count = rowSums(Df_Cond_Plants.pa[,c(21,27)]) # Excludes Unsp. Pulses
  )
)

#Simplify categories
Legumes$Site_Type <- str_replace(Legumes$Site_Type, "Religious, monastery", "Religious")    
Legumes$Site_Type <- str_replace(Legumes$Site_Type, "Castle", "Fortified")    
Legumes$Site_Type <- str_replace(Legumes$Site_Type, "Castrum", "Fortified")  

# Convert to list
Legumes.list <- list(
    Leg_Found = Legumes$Seed_Count,
    Leg_Tot = 7,
    TC_ID = as.factor((interaction(Legumes$Chronology, 
                                   Legumes$Site_Type)))
   )

Legumes.list$TC_ID <- droplevels(Legumes.list$TC_ID)


####################
## GRAPE 
####################

Grape <- with(
  Df_Cond_Plants.pa,
  data.frame(
  Chronology = Chronology,
  Macroregion = name_macroreg,
  Site_Type = as.factor(Type),
  Geo = Geo,
  Present = ifelse(Grape>0, 1, 0)
  )
)

#Simplify categories
Grape$Site_Type <- str_replace(Grape$Site_Type, "Religious, monastery", "Religious")    
Grape$Site_Type <- str_replace(Grape$Site_Type, "Castle", "Fortified")    
Grape$Site_Type <- str_replace(Grape$Site_Type, "Castrum", "Fortified")  

# Convert to list
Grape.list <- list(
    Grape_Found = Grape$Present,
    TC_ID = as.factor((interaction(Legumes$Chronology, 
                                   Legumes$Site_Type)))
   )

Grape.list$TC_ID <- droplevels(Grape.list$TC_ID)
  

####################
## OLIVES 
####################

Olives <- with(
  Df_Cond_Plants.pa,
  data.frame(
  Chronology = Chronology,
  Macroregion = name_macroreg,
  Site_Type = as.factor(Type),
  Geo = Geo,
  Present = ifelse(Olive>0, 1, 0)
  )
)

#Simplify categories
Olives$Site_Type <- str_replace(Olives$Site_Type, "Religious, monastery", "Religious")
Olives$Site_Type <- str_replace(Olives$Site_Type, "Castle", "Fortified")    
Olives$Site_Type <- str_replace(Olives$Site_Type, "Castrum", "Fortified")  

# Convert to list
Olive.list <- list(
    Olive_Found = Olives$Present,
    TC_ID = as.factor((interaction(Olives$Chronology, 
                                   Olives$Site_Type)))
   )

Olive.list$TC_ID <- droplevels(Olive.list$TC_ID)
  

####################
## NUTS 
####################

Nuts <- with(
  Df_Cond_Plants.pa,
  data.frame(
  Chronology = Chronology,
  Macroregion = name_macroreg,
  Site_Type = as.factor(Type),
  Geo = Geo,
  Seed_Count = rowSums(Df_Cond_Plants.pa[,c(29,31)])
  )
)

#Simplify categories
Nuts$Site_Type <- str_replace(Nuts$Site_Type, "Religious, monastery", "Religious")    
Nuts$Site_Type <- str_replace(Nuts$Site_Type, "Castle", "Fortified")    
Nuts$Site_Type <- str_replace(Nuts$Site_Type, "Castrum", "Fortified")  

# Convert to list
Nuts.list <- list(
    Nuts_Found = Nuts$Seed_Count,
    Nuts_Tot = 3,
    TC_ID = as.factor((interaction(Nuts$Chronology, 
                                   Nuts$Site_Type)))
   )

Nuts.list$TC_ID <- droplevels(Nuts.list$TC_ID)
  


```

```{r}
#| echo: true
#| code-fold: true
#| eval: false
# Eval is set to false because the models have been saved

# Create a binomial model, where N is 2 because that is the total number of noble 
# grains studied in this thesis
m_ng_type <- ulam(
   alist(
        NG_Found ~ dbinom( 2 , p ),
        logit(p) <- TypeChr[TC_ID],
        TypeChr[TC_ID] ~ dnorm(0,1)
    ), data=Noble_Grain.list , chains=4
    )

#saveRDS(m_ng_type, "stan_models/m_ng_type.rds")

# Create a binomial model, where N is 2 because that is the total number of noble 
# grains studied in this thesis
m_mg_type <- ulam(
    alist(
        MG_Found ~ dbinom( 7 , p ),
        logit(p) <- TypeChr[TC_ID],
        TypeChr[TC_ID] ~ dnorm(0,1)
    ), data=Minor_Grains_Type.list , chains=4
    )

#saveRDS(m_mg_type, "stan_models/m_mg_type.rds")


# Create a binomial model, where N is 2 because that is the total number of noble 
# grains studied in this thesis
m_leg_type <- ulam(
    alist(
        Leg_Found ~ dbinom( 7 , p ),
        logit(p) <- TypeChr[TC_ID],
        TypeChr[TC_ID] ~ dnorm(0,1)
    ), data=Legumes.list , chains=4
    )

#saveRDS(m_leg_type, "stan_models/m_leg_type.rds")

# Create a binomial model, where N is 2 because that is the total number of noble 
# grains studied in this thesis
m_grape_type <- ulam(
    alist(
        Grape_Found ~ dbinom( 1 , p ), # Basically a bernoulli dist
        logit(p) <- TypeChr[TC_ID],
        TypeChr[TC_ID] ~ dnorm(0,2)
    ), data=Grape.list , chains=4
    )

#saveRDS(m_grape_type, "stan_models/m_grape_type.rds")

# Create a binomial model, where N is 2 because that is the total number of noble 
# grains studied in this thesis
m_olives_type <- ulam(
    alist(
        Olive_Found ~ dbinom( 1 , p ), # Basically a bernoulli dist
        logit(p) <- TypeChr[TC_ID],
        TypeChr[TC_ID] ~ dnorm(0,2)
    ), data=Olive.list , chains=4
    )

#saveRDS(m_olives_type, "stan_models/m_olives_type.rds")

# NUTS
# Create a binomial model, where N is 3 because that is the total number of nuts
#  studied in this thesis
m_nuts_type <- ulam(
    alist(
        Nuts_Found ~ dbinom( 3 , p ),
        logit(p) <- TypeChr[TC_ID],
        TypeChr[TC_ID] ~ dnorm(0,1)
    ), data=Nuts.list , chains=4
    )

# NUTS
# Create a binomial model, where N is 3 because that is the total number of noble 
# grains studied in this thesis
m_berries_type <- ulam(
    alist(
        Berries_Found ~ dbinom( 9 , p ),
        logit(p) <- TypeChr[TC_ID],
        TypeChr[TC_ID] ~ dnorm(0,1)
    ), data=Berries_type.list , chains=4
    )

#saveRDS(m_berries_type, "stan_models/m_berries_type.rds")

```

```{r}
#| echo: false
#| include: false

## NOBLE GRAINS

m_ng_type<-readRDS("stan_models/m_ng_type.rds")

m_ng_type.df <- m_ng_type %>%
  spread_draws(TypeChr[TC_ID])

m_ng_type.df %>% mutate(Type = case_when(
  TC_ID >= 1 &  TC_ID <= 3 ~ "Fortified",
  TC_ID >= 4 & TC_ID <= 6 ~ "Necropolis",
  TC_ID >= 7 & TC_ID <= 9  ~ "Religious",
  TC_ID >= 10 & TC_ID<=13 ~ "Rural",
  TC_ID >=14 & TC_ID<=16 ~ "Villa",
  TC_ID >=17 & TC_ID <=20 ~ "Urban"
))  %>% mutate(Chrono = case_when(
    TC_ID == 6 | TC_ID==9 |  TC_ID == 13 | TC_ID==16 | TC_ID==20  ~ "R",
    TC_ID == 2 |  TC_ID == 5 | TC_ID == 11 | TC_ID==15 | TC_ID==18  ~ "LR",
    TC_ID == 1 |  TC_ID == 4 | TC_ID == 7 | TC_ID==10 | TC_ID==14 | TC_ID==17  ~ "EMA",
    TC_ID == 3 |  TC_ID == 8 | TC_ID == 12 | TC_ID==19 ~ "Ma"
)) -> m_ng_type.df

m_ng_type.plot <- ggplot(
  data = m_ng_type.df,
  aes(
    y = (factor(Chrono, levels = c("R", "LR", "EMA", "Ma"))),
    x = inv_logit(TypeChr)
  )
) +
  stat_interval(
    .width = c(.50, .80, .95, .99),
    linewidth = 3
  ) +
  geom_vline(
    xintercept = inv_logit(mean(m_ng_type.df$TypeChr)),
    linetype = "twodash",
    linewidth = 0.45,
    color = "red4"
  ) +
  scale_color_brewer(palette = "Greys") +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 10),
    sec.axis = dup_axis()
  ) +
  theme_tidybayes() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(
      fill = "grey",
      linewidth = 0.1,
      colour = "black"
    ),
    panel.background = element_rect(fill = "grey90"),
    panel.border = element_rect(fill = NA, linewidth = 0.1),
    panel.grid.major.x = element_line(
      color = "grey50",
      linewidth = 0.25,
      linetype = 2
    ),
    axis.title.x.top = element_blank()
  ) +
  facet_grid(Type ~ ., scales = "free") +
  labs(
    title = "Noble Grains",
    x = "Probability",
    y = "Chronology",
    color = "Credible interval",
    caption = "The dashed line indicates the mean for each site type and chronology."
  )


```

```{r}
#| echo: false
#| include: false


## MINOR GRAINS
m_mg_type <- readRDS("stan_models/m_mg_type.rds")

m_mg_type.df <- m_mg_type %>%
  spread_draws(TypeChr[TC_ID])

m_mg_type.df %>% mutate(Type = case_when(
  TC_ID >= 1 &  TC_ID <= 3 ~ "Fortified",
  TC_ID >= 4 & TC_ID <= 6 ~ "Necropolis",
  TC_ID >= 7 & TC_ID <= 9  ~ "Religious",
  TC_ID >= 10 & TC_ID<=13 ~ "Rural",
  TC_ID >=14 & TC_ID<=16 ~ "Villa",
  TC_ID >=17 & TC_ID <=20 ~ "Urban"
))  %>% mutate(Chrono = case_when(
    TC_ID == 6 | TC_ID==9 |  TC_ID == 13 | TC_ID==16 | TC_ID==20  ~ "R",
    TC_ID == 2 |  TC_ID == 5 | TC_ID == 11 | TC_ID==15 | TC_ID==18  ~ "LR",
    TC_ID == 1 |  TC_ID == 4 | TC_ID == 7 | TC_ID==10 | TC_ID==14 | TC_ID==17  ~ "EMA",
    TC_ID == 3 |  TC_ID == 8 | TC_ID == 12 | TC_ID==19 ~ "Ma"
)) -> m_mg_type.df


m_mg_type.plot <- ggplot(
  data = m_mg_type.df,
  aes(
    y = (factor(Chrono, levels = c("R", "LR", "EMA", "Ma"))),
    x = inv_logit(TypeChr)
  )
) +
  stat_interval(
    .width = c(.50, .80, .95, .99),
    linewidth = 3
  ) +
  geom_vline(
    xintercept = inv_logit(mean(m_mg_type.df$TypeChr)),
    linetype = "twodash",
    linewidth = 0.45,
    color = "red4"
  ) +
  scale_color_brewer(palette = "Greys") +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 10),
    sec.axis = dup_axis()
  ) +
  theme_tidybayes() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(
      fill = "grey",
      linewidth = 0.1,
      colour = "black"
    ),
    panel.background = element_rect(fill = "grey90"),
    panel.border = element_rect(fill = NA, linewidth = 0.1),
    panel.grid.major.x = element_line(
      color = "grey50",
      linewidth = 0.25,
      linetype = 2
    ),
    axis.title.x.top = element_blank()
  ) +
  facet_grid(Type ~ ., scales = "free") +
  labs(
    title = "Minor Grains",
    x = "Probability",
    y = "Chronology",
    color = "Credible interval",
    caption = "The dashed line indicates the mean for each site type and chronology."
  )


```

```{r}
#| echo: false
#| include: false

## LEGUMES
m_leg_type <- readRDS("stan_models/m_leg_type.rds")

m_leg_type.df <- m_leg_type %>%
  spread_draws(TypeChr[TC_ID])

m_leg_type.df %>% mutate(Type = case_when(
  TC_ID >= 1 &  TC_ID <= 3 ~ "Fortified",
  TC_ID >= 4 & TC_ID <= 6 ~ "Necropolis",
  TC_ID >= 7 & TC_ID <= 9  ~ "Religious",
  TC_ID >= 10 & TC_ID<=13 ~ "Rural",
  TC_ID >=14 & TC_ID<=16 ~ "Villa",
  TC_ID >=17 & TC_ID <=20 ~ "Urban"
))  %>% mutate(Chrono = case_when(
    TC_ID == 6 | TC_ID==9 |  TC_ID == 13 | TC_ID==16 | TC_ID==20  ~ "R",
    TC_ID == 2 |  TC_ID == 5 | TC_ID == 11 | TC_ID==15 | TC_ID==18  ~ "LR",
    TC_ID == 1 |  TC_ID == 4 | TC_ID == 7 | TC_ID==10 | TC_ID==14 | TC_ID==17  ~ "EMA",
    TC_ID == 3 |  TC_ID == 8 | TC_ID == 12 | TC_ID==19 ~ "Ma"
)) -> m_leg_type.df

m_leg_type.plot <- ggplot(
  data = m_leg_type.df,
  aes(
    y = (factor(Chrono, levels = c("R", "LR", "EMA", "Ma"))),
    x = inv_logit(TypeChr)
  )
) +
  stat_interval(
    .width = c(.50, .80, .95, .99),
    linewidth = 3
  ) +
  geom_vline(
    xintercept = inv_logit(mean(m_leg_type.df$TypeChr)),
    linetype = "twodash",
    linewidth = 0.45,
    color = "red4"
  ) +
  scale_color_brewer(palette = "Greys") +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 10),
    sec.axis = dup_axis()
  ) +
  theme_tidybayes() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(
      fill = "grey",
      linewidth = 0.1,
      colour = "black"
    ),
    panel.background = element_rect(fill = "grey90"),
    panel.border = element_rect(fill = NA, linewidth = 0.1),
    panel.grid.major.x = element_line(
      color = "grey50",
      linewidth = 0.25,
      linetype = 2
    ),
    axis.title.x.top = element_blank()
  ) +
  facet_grid(Type ~ ., scales = "free") +
  labs(
    title = "Legumes",
    x = "Probability",
    y = "Chronology",
    color = "Credible interval",
    caption = "The dashed line indicates the mean for each site type and chronology."
  )

```

```{r}
#| echo: false
#| include: false

m_grape_type <- readRDS("stan_models/m_grape_type.rds")

m_grape_type.df <- m_grape_type %>%
  spread_draws(TypeChr[TC_ID])

m_grape_type.df %>% mutate(Type = case_when(
  TC_ID >= 1 &  TC_ID <= 3 ~ "Fortified",
  TC_ID >= 4 & TC_ID <= 6 ~ "Necropolis",
  TC_ID >= 7 & TC_ID <= 9  ~ "Religious",
  TC_ID >= 10 & TC_ID<=13 ~ "Rural",
  TC_ID >=14 & TC_ID<=16 ~ "Villa",
  TC_ID >=17 & TC_ID <=20 ~ "Urban"
))  %>% mutate(Chrono = case_when(
    TC_ID == 6 | TC_ID==9 |  TC_ID == 13 | TC_ID==16 | TC_ID==20  ~ "R",
    TC_ID == 2 |  TC_ID == 5 | TC_ID == 11 | TC_ID==15 | TC_ID==18  ~ "LR",
    TC_ID == 1 |  TC_ID == 4 | TC_ID == 7 | TC_ID==10 | TC_ID==14 | TC_ID==17  ~ "EMA",
    TC_ID == 3 |  TC_ID == 8 | TC_ID == 12 | TC_ID==19 ~ "Ma"
)) -> m_grape_type.df

m_grape_type.plot <- ggplot(
  data = m_grape_type.df,
  aes(
    y = (factor(Chrono, levels = c("R", "LR", "EMA", "Ma"))),
    x = inv_logit(TypeChr)
  )
) +
  stat_interval(
    .width = c(.50, .80, .95, .99),
    linewidth = 3
  ) +
  geom_vline(
    xintercept = inv_logit(mean(m_grape_type.df$TypeChr)),
    linetype = "twodash",
    linewidth = 0.45,
    color = "red4"
  ) +
  scale_color_brewer(palette = "Greys") +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 10),
    sec.axis = dup_axis()
  ) +
  theme_tidybayes() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(
      fill = "grey",
      linewidth = 0.1,
      colour = "black"
    ),
    panel.background = element_rect(fill = "grey90"),
    panel.border = element_rect(fill = NA, linewidth = 0.1),
    panel.grid.major.x = element_line(
      color = "grey50",
      linewidth = 0.25,
      linetype = 2
    ),
    axis.title.x.top = element_blank()
  ) +
  facet_grid(Type ~ ., scales = "free") +
  labs(
    title = "Grape",
    x = "Probability",
    y = "Chronology",
    color = "Credible interval",
    caption = "The dashed line indicates the mean for each site type and chronology."
  )


```

```{r}
#| echo: false
#| include: false

## OLIVES

# Why is it predicting means where there is no data? 

m_olives_type <- readRDS("stan_models/m_olives_type.rds")

m_olives_type.df <- m_olives_type %>%
  spread_draws(TypeChr[TC_ID])

m_olives_type.df %>% mutate(Type = case_when(
  TC_ID >= 1 &  TC_ID <= 3 ~ "Fortified",
  TC_ID >= 4 & TC_ID <= 6 ~ "Necropolis",
  TC_ID >= 7 & TC_ID <= 9  ~ "Religious",
  TC_ID >= 10 & TC_ID<=13 ~ "Rural",
  TC_ID >=14 & TC_ID<=16 ~ "Villa",
  TC_ID >=17 & TC_ID <=20 ~ "Urban"
))  %>% mutate(Chrono = case_when(
    TC_ID == 6 | TC_ID==9 |  TC_ID == 13 | TC_ID==16 | TC_ID==20  ~ "R",
    TC_ID == 2 |  TC_ID == 5 | TC_ID == 11 | TC_ID==15 | TC_ID==18  ~ "LR",
    TC_ID == 1 |  TC_ID == 4 | TC_ID == 7 | TC_ID==10 | TC_ID==14 | TC_ID==17  ~ "EMA",
    TC_ID == 3 |  TC_ID == 8 | TC_ID == 12 | TC_ID==19 ~ "Ma"
)) -> m_olives_type.df

m_olives_type.plot <- ggplot(
  data = m_olives_type.df,
  aes(
    y = (factor(Chrono, levels = c("R", "LR", "EMA", "Ma"))),
    x = inv_logit(TypeChr)
  )
) +
  stat_interval(
    .width = c(.50, .80, .95, .99),
    linewidth = 3
  ) +
  geom_vline(
    xintercept = inv_logit(mean(m_olives_type.df$TypeChr)),
    linetype = "twodash",
    linewidth = 0.45,
    color = "red4"
  ) +
  scale_color_brewer(palette = "Greys") +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 10),
    sec.axis = dup_axis()
  ) +
  theme_tidybayes() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(
      fill = "grey",
      linewidth = 0.1,
      colour = "black"
    ),
    panel.background = element_rect(fill = "grey90"),
    panel.border = element_rect(fill = NA, linewidth = 0.1),
    panel.grid.major.x = element_line(
      color = "grey50",
      linewidth = 0.25,
      linetype = 2
    ),
    axis.title.x.top = element_blank()
  ) +
  facet_grid(Type ~ ., scales = "free") +
  labs(
    title = "Olive",
    x = "Probability",
    y = "Chronology",
    color = "Credible interval",
    caption = "The dashed line indicates the mean for each site type and chronology."
  )


```

```{r}
#| echo: false
#| include: false

## NUTS
m_nuts_type <- readRDS("stan_models/m_nuts_type.rds")

m_nuts_type.df <- m_nuts_type %>%
  spread_draws(TypeChr[TC_ID])

m_nuts_type.df %>% mutate(Type = case_when(
  TC_ID >= 1 &  TC_ID <= 3 ~ "Fortified",
  TC_ID >= 4 & TC_ID <= 6 ~ "Necropolis",
  TC_ID >= 7 & TC_ID <= 9  ~ "Religious",
  TC_ID >= 10 & TC_ID<=13 ~ "Rural",
  TC_ID >=14 & TC_ID<=16 ~ "Villa",
  TC_ID >=17 & TC_ID <=20 ~ "Urban"
))  %>% mutate(Chrono = case_when(
    TC_ID == 6 | TC_ID==9 |  TC_ID == 13 | TC_ID==16 | TC_ID==20  ~ "R",
    TC_ID == 2 |  TC_ID == 5 | TC_ID == 11 | TC_ID==15 | TC_ID==18  ~ "LR",
    TC_ID == 1 |  TC_ID == 4 | TC_ID == 7 | TC_ID==10 | TC_ID==14 | TC_ID==17  ~ "EMA",
    TC_ID == 3 |  TC_ID == 8 | TC_ID == 12 | TC_ID==19 ~ "Ma"
)) -> m_nuts_type.df


m_nuts_type.plot <- ggplot(
  data = m_nuts_type.df,
  aes(
    y = (factor(Chrono, levels = c("R", "LR", "EMA", "Ma"))),
    x = inv_logit(TypeChr)
  )
) +
  stat_interval(
    .width = c(.50, .80, .95, .99),
    linewidth = 3
  ) +
  geom_vline(
    xintercept = inv_logit(mean(m_nuts_type.df$TypeChr)),
    linetype = "twodash",
    linewidth = 0.45,
    color = "red4"
  ) +
  scale_color_brewer(palette = "Greys") +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 10),
    sec.axis = dup_axis()
  ) +
  theme_tidybayes() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(
      fill = "grey",
      linewidth = 0.1,
      colour = "black"
    ),
    panel.background = element_rect(fill = "grey90"),
    panel.border = element_rect(fill = NA, linewidth = 0.1),
    panel.grid.major.x = element_line(
      color = "grey50",
      linewidth = 0.25,
      linetype = 2
    ),
    axis.title.x.top = element_blank()
  ) +
  facet_grid(Type ~ ., scales = "free") +
  labs(
    title = "Nuts",
    x = "Probability",
    y = "Chronology",
    color = "Credible interval",
    caption = "The dashed line indicates the mean for each site type and chronology."
  )

```

```{r}
#| echo: false
#| include: false

## Stone fruits
#Berries = c("Blackberry", "Elderberry", "Blackthorn", "Wild.Cherry", "Corn..Cherry", "Sorbus.sp.", "Strawberry"),
#  Stone.fruits = c("Peach", "Plum", "Sour.Cherry", "Unsp..Prunus"),


DomFruit <- with(
  Df_Cond_Plants.pa,
  data.frame(
  Chronology = Chronology,
  Macroregion = name_macroreg,
  Site_Type = as.factor(Type),
  Geo = Geo,
  Seed_Count = rowSums(Df_Cond_Plants.pa[,c(32,33,35:38,50)])
  )
)

#Simplify categories
DomFruit$Site_Type <- str_replace(DomFruit$Site_Type, "Religious, monastery", "Religious")
DomFruit$Site_Type <- str_replace(DomFruit$Site_Type, "Castle", "Fortified")    
DomFruit$Site_Type <- str_replace(DomFruit$Site_Type, "Castrum", "Fortified")  

# Convert to list
DomFruit.list <- list(
    DF_Found = DomFruit$Seed_Count,
    DF_Tot = 7,
    TC_ID = as.factor((interaction(DomFruit$Chronology, 
                                   DomFruit$Site_Type)))
   )

DomFruit.list$TC_ID <- droplevels(DomFruit.list$TC_ID)

### BERRIES

Berries <- with(
  Df_Cond_Plants.pa,
  data.frame(
  Chronology = Chronology,
  Macroregion = name_macroreg,
  Site_Type = as.factor(Type),
  Geo = Geo,
  Seed_Count = rowSums(Df_Cond_Plants.pa[,c(41:49)])
  )
)

#Simplify categories
Berries$Site_Type <- str_replace(Berries$Site_Type, "Religious, monastery", "Religious")
Berries$Site_Type <- str_replace(Berries$Site_Type, "Castle", "Fortified")    
Berries$Site_Type <- str_replace(Berries$Site_Type, "Castrum", "Fortified")  

# Convert to list
Berries_type.list <- list(
    Berries_Found = Berries$Seed_Count,
    DF_Tot = 9,
    TC_ID = as.factor((interaction(Berries$Chronology, 
                                   Berries$Site_Type)))
   )

Berries_type.list$TC_ID <- droplevels(Berries_type.list$TC_ID)

Berries_reg.list <- list(
    Berries_Found = Berries$Seed_Count,
    DF_Tot = 9,
    RC_ID = as.factor((interaction(Berries$Chronology, 
                                   Berries$Macroregion)))
   )

Berries_reg.list$RC_ID <- droplevels(Berries_reg.list$RC_ID)

```

```{r}
#| eval: false
#| echo: false

####################
## Domestic FRUITS 
####################

# Create a binomial model, where N is 3 because that is the total number of noble 
# grains studied in this thesis
m_dfruit_type <- ulam(
    alist(
        DF_Found ~ dbinom( 7 , p ),
        logit(p) <- TypeChr[TC_ID],
        TypeChr[TC_ID] ~ dnorm(0,1)
    ), data=DomFruit.list , chains=4
    )

saveRDS(m_dfruit_type, "stan_models/m_dfruit_type.rds")
```

```{r}
#| echo: false

m_dfruit_type <- readRDS("stan_models/m_dfruit_type.rds")

m_dfruit_type.df <- m_dfruit_type %>%
  spread_draws(TypeChr[TC_ID])

m_dfruit_type.df %>% mutate(Type = case_when(
  TC_ID >= 1 &  TC_ID <= 3 ~ "Fortified",
  TC_ID >= 4 & TC_ID <= 6 ~ "Necropolis",
  TC_ID >= 7 & TC_ID <= 9  ~ "Religious",
  TC_ID >= 10 & TC_ID<=13 ~ "Rural",
  TC_ID >=14 & TC_ID<=16 ~ "Villa",
  TC_ID >=17 & TC_ID <=20 ~ "Urban"
))  %>% mutate(Chrono = case_when(
  TC_ID == 6 | TC_ID==9 |  TC_ID == 13 | TC_ID==16 | TC_ID==20  ~ "R",
  TC_ID == 2 |  TC_ID == 5 | TC_ID == 11 | TC_ID==15 | TC_ID==18  ~ "LR",
  TC_ID == 1 |  TC_ID == 4 | TC_ID == 7 | TC_ID==10 | TC_ID==14 | TC_ID==17  ~ "EMA",
  TC_ID == 3 |  TC_ID == 8 | TC_ID == 12 | TC_ID==19 ~ "Ma"
)) -> m_dfruit_type.df


m_dfruit_type.plot <- ggplot(
  data = m_dfruit_type.df,
  aes(
    y = (factor(Chrono, levels = c("R", "LR", "EMA", "Ma"))),
    x = inv_logit(TypeChr)
  )
) +
  stat_interval(
    .width = c(.50, .80, .95, .99),
    linewidth = 3
  ) +
  geom_vline(
    xintercept = inv_logit(mean(m_dfruit_type.df$TypeChr)),
    linetype = "twodash",
    linewidth = 0.45,
    color = "red4"
  ) +
  scale_color_brewer(palette = "Greys") +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 10),
    sec.axis = dup_axis()
  ) +
  theme_tidybayes() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(
      fill = "grey",
      linewidth = 0.1,
      colour = "black"
    ),
    panel.background = element_rect(fill = "grey90"),
    panel.border = element_rect(fill = NA, linewidth = 0.1),
    panel.grid.major.x = element_line(
      color = "grey50",
      linewidth = 0.25,
      linetype = 2
    ),
    axis.title.x.top = element_blank()
  ) +
  facet_grid(Type ~ ., scales = "free") +
  labs(
    title = "Domestic Fruits",
    x = "Probability",
    y = "Chronology",
    color = "Credible interval",
    caption = "The dashed line indicates the mean for each site type and chronology."
  )


m_berries_type <- readRDS("stan_models/m_berries_type.rds")

m_berries_type.df <- m_berries_type %>%
  spread_draws(TypeChr[TC_ID])

m_berries_type.df %>% mutate(Type = case_when(
  TC_ID >= 1 &  TC_ID <= 3 ~ "Fortified",
  TC_ID >= 4 & TC_ID <= 6 ~ "Necropolis",
  TC_ID >= 7 & TC_ID <= 9  ~ "Religious",
  TC_ID >= 10 & TC_ID<=13 ~ "Rural",
  TC_ID >=14 & TC_ID<=16 ~ "Villa",
  TC_ID >=17 & TC_ID <=20 ~ "Urban"
))  %>% mutate(Chrono = case_when(
  TC_ID == 6 | TC_ID==9 |  TC_ID == 13 | TC_ID==16 | TC_ID==20  ~ "R",
  TC_ID == 2 |  TC_ID == 5 | TC_ID == 11 | TC_ID==15 | TC_ID==18  ~ "LR",
  TC_ID == 1 |  TC_ID == 4 | TC_ID == 7 | TC_ID==10 | TC_ID==14 | TC_ID==17  ~ "EMA",
  TC_ID == 3 |  TC_ID == 8 | TC_ID == 12 | TC_ID==19 ~ "Ma"
)) -> m_berries_type.df


m_berries_type.plot <- ggplot(
  data = m_berries_type.df,
  aes(
    y = (factor(Chrono, levels = c("R", "LR", "EMA", "Ma"))),
    x = inv_logit(TypeChr)
  )
) +
  stat_interval(
    .width = c(.50, .80, .95, .99),
    linewidth = 3
  ) +
  geom_vline(
    xintercept = inv_logit(mean(m_berries_type.df$TypeChr)),
    linetype = "twodash",
    linewidth = 0.45,
    color = "red4"
  ) +
  scale_color_brewer(palette = "Greys") +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 10),
    sec.axis = dup_axis()
  ) +
  theme_tidybayes() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(
      fill = "grey",
      linewidth = 0.1,
      colour = "black"
    ),
    panel.background = element_rect(fill = "grey90"),
    panel.border = element_rect(fill = NA, linewidth = 0.1),
    panel.grid.major.x = element_line(
      color = "grey50",
      linewidth = 0.25,
      linetype = 2
    ),
    axis.title.x.top = element_blank()
  ) +
  facet_grid(Type ~ ., scales = "free") +
  labs(
    title = "Berries",
    x = "Probability",
    y = "Chronology",
    color = "Credible interval",
    caption = "The dashed line indicates the mean for each site type and chronology."
  )

```

```{r}
#| echo: false
#| warning: false
#| label: fig-bayes-glm-site-type
#| fig-cap: "Bayesian probability estimates for main plant taxa, divided by site type and chronology. The prediction estimates are produced with a GLM and weakly informative priors. Credibility intervals are provided."
#| fig-subcap: 
#|  - "Noble grains (wheat and barley)."
#|  - "Minor grains."
#|  - "Legumes."
#|  - "Grape."
#|  - "Olives. The model is predicting (conservative) values for fortified sites, in which no olive was found."
#|  - "Nuts."
#|  - "Domestic fruits."
#|  - "Berries."
#| layout-nrow: 8
#| layout-ncol: 1
#| fig-height: 10

m_ng_type.plot
m_mg_type.plot
m_leg_type.plot
m_grape_type.plot
m_olives_type.plot
m_nuts_type.plot
m_dfruit_type.plot
m_berries_type.plot

```

### Richness and diversity

```{r}
#| echo: false


# Remove NAs
Df_Cond_Plants[is.na(Df_Cond_Plants)] <-0 

# Generate a dataframe with the relative proportions and round the results
Df_Cond_Plants_Rel <- decostand(Df_Cond_Plants[11:50], method = "total")
Df_Cond_Plants_Rel <- round(Df_Cond_Plants_Rel, digits=2)

Df_Cond_Plants_RichnDiv <- data.frame("Geo" = Df_Cond_Plants$Geo,
                                                 "Chronology" = Df_Cond_Plants$Chronology,
                                                 "Type"= Df_Cond_Plants$Type, 
                                                 "Macroregion" = Df_Cond_Plants$name_macroreg,
                                                 "Specnumber" = specnumber(Df_Cond_Plants_Rel), #All plants
                                                 "Shannon Div" = diversity(Df_Cond_Plants_Rel),
                                                 Df_Cond_Plants_Rel
)

#Simplify categories
Df_Cond_Plants_RichnDiv$Type <- str_replace(Df_Cond_Plants_RichnDiv$Type, "Religious, monastery", "Religious")    
Df_Cond_Plants_RichnDiv$Type <- str_replace(Df_Cond_Plants_RichnDiv$Type, "Castle", "Fortified")    
Df_Cond_Plants_RichnDiv$Type <- str_replace(Df_Cond_Plants_RichnDiv$Type, "Castrum", "Fortified")  

# Let's plot the diversity by type

# Creating the dataframes for R and EMA age 
Df_Cond_Plants_RichnDiv.R <- filter(Df_Cond_Plants_RichnDiv, Chronology == "R")
Df_Cond_Plants_RichnDiv.LR <- filter(Df_Cond_Plants_RichnDiv, Chronology == "LR")
Df_Cond_Plants_RichnDiv.EMA <- filter(Df_Cond_Plants_RichnDiv, Chronology == "EMA")
Df_Cond_Plants_RichnDiv.Ma <- filter(Df_Cond_Plants_RichnDiv, Chronology == "Ma")


##########################################
## CORRECTION FOR INCOMPLETE SAMPLES
##########################################

#ROMAN 
# Need to filter out sites that did not provide legumes or fruits at all
# Need to compare these rows to see if there are some indexes in common 
CompleteSamples.R <- data.frame(
Cereals = (rowSums(Df_Cond_Plants_RichnDiv.R[,c(7:16)]) > 0),
Legumes = (rowSums(Df_Cond_Plants_RichnDiv.R[,c(17:30)]) > 0),
Fruits = (rowSums(Df_Cond_Plants_RichnDiv.R[,c(31:46)]) > 0)
)

IncompleteSamplesIndexes.R <- which(rowSums(CompleteSamples.R) < 3)

# LR 
CompleteSamples.LR <- data.frame(
Cereals = (rowSums(Df_Cond_Plants_RichnDiv.LR[,c(7:16)]) > 0),
Legumes = (rowSums(Df_Cond_Plants_RichnDiv.LR[,c(17:30)]) > 0),
Fruits = (rowSums(Df_Cond_Plants_RichnDiv.LR[,c(31:46)]) > 0)
)

IncompleteSamplesIndexes.LR <- which(rowSums(CompleteSamples.LR) < 3)

# EMA 
CompleteSamples.EMA <- data.frame(
Cereals = (rowSums(Df_Cond_Plants_RichnDiv.EMA[,c(7:16)]) > 0),
Legumes = (rowSums(Df_Cond_Plants_RichnDiv.EMA[,c(17:30)]) > 0),
Fruits = (rowSums(Df_Cond_Plants_RichnDiv.EMA[,c(31:46)]) > 0)
)

IncompleteSamplesIndexes.EMA <- which(rowSums(CompleteSamples.EMA) < 3)

# Ma 
CompleteSamples.Ma <- data.frame(
Cereals = (rowSums(Df_Cond_Plants_RichnDiv.Ma[,c(7:16)]) > 0),
Legumes = (rowSums(Df_Cond_Plants_RichnDiv.Ma[,c(17:30)]) > 0),
Fruits = (rowSums(Df_Cond_Plants_RichnDiv.Ma[,c(31:46)]) > 0)
)

IncompleteSamplesIndexes.Ma <- which(rowSums(CompleteSamples.Ma) < 3)


##########################################
## END OF CORRECTION
##########################################

plot_RichnessType.R <- ggplot(Df_Cond_Plants_RichnDiv.R, aes(x = Type, y = Specnumber, fill = Type)) +
  geom_violin(trim=FALSE) + 
  geom_boxplot(width=0.1, fill="white")+
  geom_jitter(alpha=0.3)+
  scale_x_discrete(labels = c("Necropolis \n (n = 13)", "Religious \n (n = 3)", "Rural \n (n=12)", "Villa \n (n=4)", "Urban \n (n=44)" )) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 11, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = "Site type",
       y = "Number of species per site",
       title = "R - Plant richness")

plot_DiversityType.R <- ggplot(Df_Cond_Plants_RichnDiv.R, aes(x = Type, y = Shannon.Div, fill = Type)) +
  geom_violin(trim=FALSE) + 
  geom_boxplot(width=0.1, fill="white")+
  geom_jitter(alpha=0.3)+
  scale_x_discrete(labels = c("Necropolis \n (n = 13)", "Religious \n (n = 3)", "Rural \n (n=12)", "Villa \n (n=4)", "Urban \n (n=44)" )) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 11, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = "Site type",
       y = "Site diversity",
       title = "R - Shannon diversity")

plot_RichnessType.LR <- ggplot(Df_Cond_Plants_RichnDiv.LR, aes(x = Type, y = Specnumber, fill = Type)) +
  geom_violin(trim=FALSE) + 
  geom_boxplot(width=0.1, fill="white")+
  geom_jitter(alpha=0.3)+
  scale_x_discrete(labels = c("Fortified \n (n=3)", "Necropolis \n (n = 10)", "Rural \n (n=14)", "Villa \n (n=11)", "Urban \n (n=24)" )) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 11, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = "Site type",
       y = "Number of species per site",
       title = "LR - Plant richness")

plot_DiversityType.LR <- ggplot(Df_Cond_Plants_RichnDiv.LR, aes(x = Type, y = Shannon.Div, fill = Type)) +
  geom_violin(trim=FALSE) + 
  geom_boxplot(width=0.1, fill="white")+
  geom_jitter(alpha=0.3)+
  scale_x_discrete(labels = c("Fortified \n (n=3)", "Necropolis \n (n = 10)", "Rural \n (n=14)", "Villa \n (n=11)", "Urban \n (n=24)" )) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 11, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = "Site type",
       y = "Site diversity",
       title = "LR - Shannon diversity")

plot_RichnessType.EMA <- ggplot(Df_Cond_Plants_RichnDiv.EMA, aes(x = Type, y = Specnumber, fill = Type)) +
  geom_violin(trim=FALSE) + 
  geom_boxplot(width=0.1, fill="white")+
  geom_jitter(alpha=0.3)+
  scale_x_discrete(labels = c("Fortified \n (n = 7)", "Necropolis \n (n = 1)", "Religious \n (n=2)", "Rural \n (n=18)", "Villa \n (n=11)", "Urban \n (n=25)" )) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 11, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = "Site type",
       y = "Number of species per site",
       title = "EMA - Plant richness")

plot_DiversityType.EMA <- ggplot(Df_Cond_Plants_RichnDiv.EMA, aes(x = Type, y = Shannon.Div, fill = Type)) +
  geom_violin(trim=FALSE) + 
  geom_boxplot(width=0.1, fill="white")+
  geom_jitter(alpha=0.3)+
  scale_x_discrete(labels = c("Fortified \n (n = 7)", "Necropolis \n (n = 1)", "Religious \n (n=2)", "Rural \n (n=18)", "Villa \n (n=11)", "Urban \n (n=25)" )) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 11, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = "Site type",
       y = "Site diversity",
       title = "EMA - Shannon Diversity")


```

```{r}
#| echo: false
#| eval: false

# Check mean of plant richness
# [RICHNESS] Urban Roman: 6.5 | Urban LR: 5.9 | Urban EMA: 7.2 | Urban Ma: 9
# [CORRECTED RICHNESS]: UrbR: 8.45 | UrbLR: 7.2 | UrbEMA: 9.625 | UrbMa: 17.5
# [DIVERSITY] Urban Roman: 1.02 | Urban LR: 1.20 | Urban EMA: 1.45 |Urban Ma: 1.39

mean((filter(Df_Cond_Plants_RichnDiv.R, Type=="Urban"))$Specnumber)
mean((filter(Df_Cond_Plants_RichnDiv.R[-IncompleteSamplesIndexes.R,], Type=="Urban"))$Specnumber)
mean((filter(Df_Cond_Plants_RichnDiv.LR, Type=="Urban"))$Specnumber)
mean((filter(Df_Cond_Plants_RichnDiv.LR[-IncompleteSamplesIndexes.LR,], Type=="Urban"))$Specnumber)
mean((filter(Df_Cond_Plants_RichnDiv.EMA, Type=="Urban"))$Shannon.Div)
mean((filter(Df_Cond_Plants_RichnDiv.EMA[-IncompleteSamplesIndexes.EMA,], Type=="Urban"))$Shannon.Div)
mean((filter(Df_Cond_Plants_RichnDiv.Ma, Type=="Urban"))$Specnumb)
mean((filter(Df_Cond_Plants_RichnDiv.Ma[-IncompleteSamplesIndexes.Ma,], Type=="Urban"))$Specnumber)


```

```{r}
#| echo: false
#| warning: false
#| layout: "[[100,100], [100,100]], [100,100]"
#| label: fig-type-richness-div
#| fig-cap: "Violin plots of plant richness and diversity in different site types. The grey dots (jitters) indicate the value for the single site, while the white boxplot shows the median and the quartile values." 
#| fig-subcap: 
#|   - "Roman plant richness" 
#|   - "Roman plant diversity."
#|   - "Late Roman plant richness" 
#|   - "Late Roman plant diversity."
#|   - "EMA plant richness" 
#|   - "EMA plant diversity."
#| fig-height: 4
#| fig-width: 6
#| fig-dpi: 600

plot_RichnessType.R
plot_DiversityType.R
plot_RichnessType.LR
plot_DiversityType.LR
plot_RichnessType.EMA
plot_DiversityType.EMA

```

#### Richness in urban contexts

The richness of plants in urban contexts has been calculated for every phase. As not every sample reported cereals, fruits/nuts and legumes, a correction has also been applied to exclude contexts that did not report the three types of plant taxa under examination. This correction has however strong disadvantages as it removes sites where the absence of certain plant categories may have been deliberate (*e.g.*, storages). However, if the figures vary considerably (especially in the case of the 11^th^ century CE which only consists of 2 samples after the correction), the general trend is similar after the correction. Plant richness in urban contexts slightly decreases after the Roman age, but early medieval cities are richer in plants when compared to the Roman and Late Roman phases. In the 11th century, plant richness increases further. These values will be discussed in the results chapter.

| Chronology | Richness | Samples | Richness\* | Samples\* |
|------------|----------|---------|------------|-----------|
| R          | 6.5      | 44      | 8.45       | 22        |
| LR         | 5.9      | 24      | 7.2        | 10        |
| EMA        | 7.2      | 25      | 9.6        | 8         |
| Ma         | 9        | 8       | 17.5       | 2         |

: Plant richness in urban contexts before and after (**\***) removing incomplete samples. {#tbl-richness-urban}

## Soil types vs Crops

see Rippon et al 2015, p. 81 as a comparison for results and Lodwick chapter in Hamerow and McKerracher 2022, p. 153

# Macroregion {style="text-align: justify;"}

## Cereals

### Richness and diversity {#sec-richness-diversity-macroregions style="text-align:justify;"}

::: callout-important
## Section in progress {style="text-align:justify;"}
:::

::: {.content-visible when-format="html"}
```{r}
#| echo: false
#| output: false
#| code-summary: "Show the code: data preparation"
#| code-overflow: wrap

# Species richness based on geographical features
# RELATIVE PROPORTIONS OF ARCHAEOBOT_VIZ QUERY EXPORT FROM THE DB
# (Condensed, without totals)


# Add more info to the dataframe
Df_Cond_Plants_Rel_Richness_Diversity <- data.frame(
                                 "Geo" = Df_Cond_Plants$Geo,
                                 "Chronology" = Df_Cond_Plants$Chronology, 
                                 "Type"= Df_Cond_Plants$Type, 
                                 "Specnumber" = specnumber(Df_Cond_Plants_Rel),
                                 "Shannon Div" = diversity(Df_Cond_Plants_Rel),
                                 Df_Cond_Plants_Rel
)

Df_Cond_Plants_Rel_withMacroregion <- data.frame("Geo" = Df_Cond_Plants$Geo,
                                                 "Chronology" = Df_Cond_Plants$Chronology,
                                                 "Type"= Df_Cond_Plants$Type, 
                                                 "Macroregion" = Df_Cond_Plants$name_macroreg,
                                                 "Specnumber" = specnumber(Df_Cond_Plants_Rel[1:10]), #Only cereals
                                                 "Shannon Div" = diversity(Df_Cond_Plants_Rel[1:10]),
                                                 Df_Cond_Plants_Rel[1:10]
)

# Let's plot the diversity by macroregion

# Creating the dataframes for R and EMA age 
# I know it's called "Plants" but it's actually just cereals
Df_Cond_Plants_Rel_withMacroregion.R <- filter(Df_Cond_Plants_Rel_withMacroregion, Chronology == "R")
Df_Cond_Plants_Rel_withMacroregion.LR <- filter(Df_Cond_Plants_Rel_withMacroregion, Chronology == "LR")
Df_Cond_Plants_Rel_withMacroregion.EMA <- filter(Df_Cond_Plants_Rel_withMacroregion, Chronology == "EMA")

```

```{r}
#| echo: false
#| output: false
#| code-summary: "Show the code: plots"
#| code-overflow: wrap

pal_RichnessvsGeo <- c("cadetblue3", "gold1",  "bisque4", "palegreen4")

plot_RichnessMacroReg.R <- ggplot(Df_Cond_Plants_Rel_withMacroregion.R, aes(x = Macroregion, y = Specnumber, fill = Macroregion)) +
  geom_violin(trim=FALSE) + 
  geom_boxplot(width=0.1, fill="white")+
  scale_fill_manual(values = pal_RichnessvsGeo) +
  geom_jitter(alpha=0.3)+
  scale_x_discrete(labels = c("Central Italy \n (n = 5)", "Northern Italy \n (n = 41)", "Southern Italy \n (n=34)")) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = "Macroregion",
       y = "Number of species per site",
       title = "R - Cereal richness")

plot_RichnessMacroReg.EMA <- ggplot(Df_Cond_Plants_Rel_withMacroregion.EMA, aes(x = Macroregion, y = Specnumber, fill = Macroregion)) +
  geom_violin(trim=FALSE) + 
  geom_boxplot(width=0.1, fill="white")+
  scale_fill_manual(values = pal_RichnessvsGeo) +
  geom_jitter(alpha=0.3)+
  scale_x_discrete(labels = c("Central Italy \n (n = 11)", "Northern Italy \n (n = 36)", "Southern Italy \n (n=17)")) +
  theme(legend.position = "none",
        plot.background = element_rect("white"),
        panel.background = element_rect("white"),
        panel.grid = element_line("grey90"),
        axis.line = element_line("gray25"),
        axis.text = element_text(size = 12, color = "gray25"),
        axis.title = element_text(color = "gray25"),
        legend.text = element_text(size = 12)) + 
  labs(x = "Macroregion",
       y = "Number of species per site",
       title = "EMA - Cereal richness")

# Check median of cereal richness per macroregion

# ROMAN NORTH
mean((filter(Df_Cond_Plants_Rel_withMacroregion.R, Macroregion=="Northern Italy"))$Specnumber)
#ROMAN SOUTH
mean((filter(Df_Cond_Plants_Rel_withMacroregion.R, Macroregion=="Southern Italy"))$Specnumber)
# EMA NORTH
mean((filter(Df_Cond_Plants_Rel_withMacroregion.EMA, Macroregion=="Northern Italy"))$Specnumber)
#EMA SOUTH
mean((filter(Df_Cond_Plants_Rel_withMacroregion.EMA, Macroregion=="Southern Italy"))$Specnumber)

```
:::

Cereals share similar presence values in Roman Northern and Southern Italian sites (@fig-cereal-richness-RvsEMA). Central Italy reports higher values, although this is based only on three sites and hence it is not reliable. During the Early Middle Ages, Central Italy again is the richest in cereals, closely followed by Northern Italy. Interestingly, Southern Italy still reports values very close to the Roman age. A full list of the Southern Italian EMA sites is reported in @tbl-southern-ema-sites.

```{r}
#| echo: false
#| layout: "[[100,100], [1]]"
#| label: fig-cereal-richness-RvsEMA
#| fig-cap: "Violin plots of cereal richness in the Italian macroregions. The grey dots (jitters) indicate the value for the single site, while the white boxplot shows the median and the quartile values." 
#| fig-subcap: 
#|   - "Roman age." 
#|   - "Early Medieval age."
#| fig-height: 4
#| fig-width: 6
#| fig-dpi: 600

plot_RichnessMacroReg.R
plot_RichnessMacroReg.EMA

```

| ID  | Site                                   | Region     | Geography    | Type         | Culture/Influence |
|------------|------------|------------|------------|------------|------------|
| 98  | S. Maria in Civit, D85                | Molise     | Hilltop      | Urban        | Lombard           |
| 107 | S. Giovanni di Ruoti, Phase 3A         | Basilicata | Mountain     | Monastery    | Lombard           |
| 107 | S. Giovanni di Ruoti, Phase 3B         | Basilicata | Mountain     | Monastery    | Lombard           |
| 198 | Salapia, area botteghe, US 2475        | Puglia     | Coast/Lagoon | Urban        | Lombard           |
| 198 | Salapia, area botteghe, US 2437        | Puglia     | Coast/Lagoon | Urban        | Lombard           |
| 199 | Salapia, area conceria, US 2054        | Puglia     | Coast/Lagoon | Urban        | Lombard           |
| 199 | Salapia, area conceria, US 2211-2217   | Puglia     | Coast/Lagoon | Urban        | Lombard           |
| 199 | Salapia, area conceria, 8th-9th c.     | Puglia     | Coast/Lagoon | Urban        | Lombard           |
| 196 | Faragola, wastepit 61                  | Puglia     | Plain        | Rural, villa | Lombard           |
| 196 | Faragola, wastepit 66                  | Puglia     | Plain        | Rural, villa | Lombard           |
| 234 | Colle Castellano, Phase 3-4            | Molise     | Hill         | Urban        | Lombard           |
| 177 | San Vincenzo al Volturno, kitchen area | Molise     | Hill         | Monastery    | Lombard           |
| 101 | Supersano, loc. Scorpo                 | Puglia     | Plain        | Rural        | Byzantine         |
| 250 | Apigliano, 9th-10th c., pits           | Puglia     | Plain        | Rural        | Byzantine         |
| 250 | Apigliano, 10th-11th c., pits          | Puglia     | Plain        | Rural        | Byzantine         |
| 196 | Faragola, granary A7                   | Puglia     | Plain        | Rural, villa | Lombard           |
| 196 | Faragola, granary A8                   | Puglia     | Plain        | Rural, villa | Lombard           |

: List of Southern Italian sites with chronology EMA {#tbl-southern-ema-sites}

### Assessing the difference {#sec-permanova-cereals}

::: callout-note
## Notes on terminology: PERMANOVA {style="text-align:justify;"}

Permutational multivariate analysis of variance (PERMANOVA) is a non-parametric multivariate statistical test used to compare group of objects. By using measure space, the null hypothesis that the centroids and dispersion of groups are identical is tested. The null hypothesis is rejected if either the centroid or the spread of the objects differs between the groups. A prior calculation of the distance between any two objects included in the experiment is used to determine whether the test is valid or not[^archaeobotany-2] [@andersonPermutationalMultivariateAnalysis2017]. In this context, the *null hypothesis* is that there is no regional difference in the cereals dataset, with cereals being evenly distributed across macroregions and chronologies.
:::

[^archaeobotany-2]: Source: Wikipedia. Change the source later

The suggestion of an Early Medieval shift in cereal farming stated in @sec-macroregional-differences-ubiquity and @sec-richness-diversity-macroregions needs statistical support. Considering that data is not unimodal and that we are dealing with presence/absence analysis, the best choice is to use a non-parametric test as PERMANOVA on the early medieval botanical dataset. Prior to performing the test, it was necessary to pre-process data by:

-   Selecting all the cereals columns of the plant remains table, keeping the categorical variables: `Macroregion`.

-   Removing empty rows (samples that only reported seeds/fruits, but not cereals).

-   Transforming the raw counts into presence/absence, using the function `decostand()` (`method=pa`) in the `R` package `vegan` [@vegan].

::: {.content-visible when-format="html"}
```{r}
#| code-fold: true
#| output: false
#| code-summary: "Show the code: Pre-processing"
#| code-overflow: wrap
# Testing the results: Regionality in the dataset? 
library(vegan)

set.seed(29)

# Pre processing: remove empty rows

# Note: The input table is the CONDENSED table without totals

# Selecting all the cereals columns of the plant remains table, keeping some categorical variables 
cer_macroreg_ubiquity_transp.tot <- Df_Cond_Plants[c(4,5,6,7,11:19)]

# Selecting all rows with data (since we selected only with cereals, and some sites only had fruits/pulses we might have empty rows)
cer_macroreg_ubiquity_transp.tot <- cer_macroreg_ubiquity_transp.tot[rowSums(cer_macroreg_ubiquity_transp.tot[5:13])>0,]

# Assigning a column name "Macroregion"
colnames(cer_macroreg_ubiquity_transp.tot)[1] = "Macroregion"

# Selecting the Chronology of interest (EMA) and excluding Central Italy
cer_macroreg_ubiquity_transp.tot<- filter(cer_macroreg_ubiquity_transp.tot, Macroregion!="Central Italy" & Chronology=="EMA")

# dividing categorical and numerical columns
cer_macroreg_ubiquity_transp.categ <- cer_macroreg_ubiquity_transp.tot[1:4]
cer_macroreg_ubiquity_transp.data <- cer_macroreg_ubiquity_transp.tot[5:13]

# Converting the numerical columns into a presence/absence matrix (using method=pa)
cer_macroreg_ubiquity_transp.dist <- decostand(cer_macroreg_ubiquity_transp.data, method="pa", na.rm=TRUE)


```
:::

After the pre-processing, it was possible to run the PERMANOVA using the function `adonis2()` in the package `vegan`. The function creates a distance matrix and computes an analysis of variance on the matrix. The method chosen to calculate the distance matrix is the `jaccard` distance. The Jaccard distance, based on the Jaccard similarity index, is a value of dissimilarity between sample sets [@kosubNoteTriangleInequality2019]. When compared to other dissimilarity indices, it is more appropriate for presence/absence analyses as it is not based on Euclidean distance.

**Results of `adonis2()`**.

::: {.content-visible when-format="html"}
```{r}
#| code-fold: true
#| output: false
#| code-summary: "Show the code: adonis2() "
#| code-overflow: wrap

cer_macroreg_ubiquity_transp.div <- adonis2(
  cer_macroreg_ubiquity_transp.dist ~ Macroregion, 
  data = cer_macroreg_ubiquity_transp.categ,
  permutations = 10000, method="jaccard"
  )

```
:::

```{r}
#| echo: false

cer_macroreg_ubiquity_transp.div

```

The calculation of PERMANOVA on the Roman dataset using the variable `Macroregion` reported a not statistically significant p-value, suggesting a degree of homogeneity. On the other hand, when applied to the early Medieval dataset, the results were highly significant (0\<p\<0.001) with a 99.99% confidence. PERMANOVA has not been calculated on the Late Roman dataset as samples from Southern Italy were too scarce. To confirm the significance of the results of PERMANOVA applied to the EMA dataset, it is necessary to check that its assumptions are met (especially since we are dealing with small groups of data). Firstly, we check the homogeneity of variances. The function `betadisper()` from the package `vegan` provides the distances of group samples from centroids. If the variation is even, the null hypothesis of no difference in dispersion between groups is accepted. To test the variation, it is possible to use the analysis of variance (ANOVA).

::: {.content-visible when-format="html"}
```{r}
#| code-fold: true
#| output: false
#| code-summary: "Show the code: betadisper()"
#| code-overflow: wrap

# We do not need to calculate the distance separately, but it will be useful later for the betadisper() function

# Distance dissimilarity matrix with the Jaccard method
cer_macroreg_ubiquity_transp.dist2 <- vegdist(cer_macroreg_ubiquity_transp.dist, method="jaccard", na.rm=TRUE)

# Betadisper: distances of group samples from centroids
cer_macroreg_ubiquity_transp.betadisper <- betadisper(cer_macroreg_ubiquity_transp.dist2, cer_macroreg_ubiquity_transp.categ$Macroregion)
```
:::

**Results of `anova()` on the betadisper.**

```{r}
#| code-fold: true
#| output: true
#| code-summary: "Show the code: ANOVA on betadisper()"
#| code-overflow: wrap

# We will see that the ANOVA's p-value is not significant meaning that group dispersions are homogenous 
#("Null hypothesis of no difference in dispersion between groups"; https://www.rdocumentation.org/packages/vegan/versions/2.4-2/topics/betadisper).

anova(cer_macroreg_ubiquity_transp.betadisper) # This should not be significant!
```

```{r}
#| echo: false
#| layout: "[[100,100], [1]]"
#| label: fig-betadisper
#| fig-cap: "Results of the betadisper() (groups dispersions) on the distance matrix calculated with the Jaccard method." 
#| fig-subcap: 
#|   - "Groups dispersions plot with confidence ellipses." 
#|   - "Boxplot showing equal distances from centroid."
#| fig-height: 4
#| fig-width: 5
#| fid-dpi: 300

# Now we can plot and see that group dispersions (distances from centroids) are very similar, but compositions seem to be very different. Let's go to adonis to test that.
plot(cer_macroreg_ubiquity_transp.betadisper, hull=FALSE, ellipse=TRUE, main="Betadisper plot")
boxplot(cer_macroreg_ubiquity_transp.betadisper)

```

The `betadisper()` graphs (@fig-betadisper) show similar distances from the centroids for the categories `Northern Italy` and `Southern Italy`. In addition, the ANOVA on the `betadisper()` shows that the separation is not significant (p-value over the significance threshold), meaning that the groups dispersions are homogeneous. We can thus be confident in the PERMANOVA results and accept the difference between the two groups of sites under investigation. In other words, the Southern and Northern Italian group of sites are different during the Early Middle Ages for what concerns cereal farming.

Running the same tests on the Roman sites failed to separate the two groups of sites, confirming that there was not a significant difference in the types of cereals cultivated during the Roman age between Northern and Southern Italy.

### Quantifying the separation: nMDS {#sec-nMDS-Cereals style="text-align:justify"}

::: callout-note
## Notes on terminology: Wasserstein metric {style="text-align:justify;"}

The [Wasserstein distance](https://en.wikipedia.org/wiki/Earth_mover%27s_distance) (or earth's mover distance) is a measure of distance between two probability distributions on a metric space.
:::

In addition to statistically testing the separation between the Northern and Southern Italian early medieval cereals dataset (@sec-permanova-cereals), it is possible to measure the distance between groups of sites both in the Roman and early Middle ages. For this task, a dimensionality reduction algorithm has been chosen: the *non-metric multidimensional scaling*, from the `R` library `vegan`. A more in-depth explanation of the algorithm can be read in @sec-met-nmds. To avoid fallacy in computations, the macroregion `Central Italy` and the chronologies `LR` (Late Roman) and `Ma` (11^th^ c. onwards) have been excluded from this test---the uneven distribution of the group of samples required a cautious approach. The nMDS has been run with a reduction to only one dimension, using KDE plots to visualize the results. Setting the dimension to one allows easier calculations of distance. In @fig-nMDS-1D-macroregion (a), it is possible to see the nMDS performed on the Roman cereals presence/absence dataset. As already pointed out, the PERMANOVA did not produce significant results for this dataset and the Wasserstein distance (calculated with the `wasserstein1d()` function in the `transport` library) is indeed shorter for the Roman dataset. For both chronologies there is an overlap in the curves, which is more considerable in the Roman age (indicating that the group of samples are more similar). The overlap for the EMA groups (@fig-nMDS-1D-macroregion, b) is probably due to the fact that the presence of the noble grains is not by itself a 'marker' of Southern Italian sites---these grains are also very common in the North. The difference is that in the South noble grains are not cultivated in conjunction with other grains. The graph for the EMA chronology shows a clearer separation of the macroregional groups, with some minor overlaps. Moreover, the graph also displays variability in the Northern Italian dataset. The variability can also be assessed from the outliers in the boxplots in @fig-boxplot-nMDS-R-EMA.

::: {.content-visible when-format="html"}
**Source code:**

```{r}
#| code-summary: "Show the code: Data preparation"
#| code-overflow: wrap
#| code-fold: true

# DATA PREPARATION #

# Creating a simple dataframe just for the purpose of the NCA
Df_Cond_Plants_simplified <- data.frame(
  "Chronology" = Df_Cond_Plants$Chronology, 
  "Type" = Df_Cond_Plants$Type,
  "Macroregion" = Df_Cond_Plants$name_macroreg,
  "Weight" = Df_Cond_Plants$weight,
  Df_Cond_Plants[11:ncol(Df_Cond_Plants)]
  )

# Joining site typology categories to simplify the output, using library stringr
library(vegan)
library(stringr)

Df_Cond_Plants_simplified$Type <- str_replace_all(
  Df_Cond_Plants_simplified$Type, c(
    "Rural site, villa" = "Rural", 
    "Rural site, mansio" = "Rural",
    "Religious, monastery" = "Religious",
    "Castle" = "Fortified",
    "Castrum" = "Fortified")
  )

# Convert to Presence (1) / Absence (0) using library vegan
Df_Cond_Plants_simplified[5:ncol(Df_Cond_Plants_simplified)] <- decostand(Df_Cond_Plants_simplified[5:ncol(Df_Cond_Plants_simplified)], method="pa")

# Creating dataframes for the chronologies R and EMA
Df_Cereals_R <- filter(Df_Cond_Plants_simplified, Chronology=="R" & Macroregion!="Central Italy")
Df_Cereals_EMA <- filter(Df_Cond_Plants_simplified, Chronology=="EMA" & Macroregion!="Central Italy")

# Selecting only cereals from both dataframes
Df_Cereals_R <- Df_Cereals_R[1:14]
Df_Cereals_R[is.na(Df_Cereals_R)] <- 0
Df_Cereals_R <- Df_Cereals_R[rowSums(Df_Cereals_R[,5:14] != 0) > 0, ] # Removing empty rows

Df_Cereals_EMA <- Df_Cereals_EMA[1:14]
Df_Cereals_EMA[is.na(Df_Cereals_EMA)] <- 0
Df_Cereals_EMA <- Df_Cereals_EMA[rowSums(Df_Cereals_EMA[,5:14] != 0) > 0, ]

```

```{r}
#| echo: false
#| output: false

library(ggpubr)
set.seed(7)

# Perform the metaMDS to 1D for Df_Cereals_R and Df_Cereals_EMA 
mds_R <- metaMDS(Df_Cereals_R[, c(5:14)], k=1, distance = "jaccard")
mds_EMA <- metaMDS(Df_Cereals_EMA[, c(5:14)], k=1, distance = "jaccard")

# Add the Macroregion data to the metaMDS results for both dataframes
mds_df_R <- data.frame(mds_R$points, Macroregion = Df_Cereals_R$Macroregion)
mds_df_EMA <- data.frame(mds_EMA$points, Macroregion = Df_Cereals_EMA$Macroregion)

# Create dataframes to calculate wasserstein distance
library(transport)

MDS_Scores_R_SI <- (filter(mds_df_R, Macroregion=="Southern Italy"))$MDS1
MDS_Scores_R_NI <- (filter(mds_df_R, Macroregion=="Northern Italy"))$MDS1

MDS_Scores_EMA_SI <- (filter(mds_df_EMA, Macroregion=="Southern Italy"))$MDS1
MDS_Scores_EMA_NI <- (filter(mds_df_EMA, Macroregion=="Northern Italy"))$MDS1

       
# Create the plots
mds_R.plot <- ggplot(
  data=mds_df_R, 
  aes(x=MDS1, group=Macroregion, color=Macroregion, fill=Macroregion)) +
    geom_density(adjust=1.5, alpha=.2) +
      geom_vline(aes(xintercept=mean(MDS_Scores_R_SI)),
            color="#90a955", linetype="dashed", size=.8)+ #South
      geom_vline(aes(xintercept=mean(MDS_Scores_R_NI)),
            color="#212529", linetype="dashed", size=.8)+ #North
  scale_fill_manual(values=c("#212529","#90a955"))+
  scale_color_manual(values=c("#212529","#90a955"))+
  theme_pubclean()+
  theme(legend.position = "none")+
    labs(
      title="Roman age",
      caption = paste0(
      "PERMANOVA: p > 0.05 \n",
      "Wasserstein distance: ", round(
      wasserstein1d(MDS_Scores_R_SI,MDS_Scores_R_NI),2
      )))

mds_EMA.plot <- ggplot(
  data=mds_df_EMA, 
  aes(x=MDS1, group=Macroregion, color=Macroregion, fill=Macroregion)
  ) +
  scale_alpha(guide = 'none')+
  geom_density(adjust=1.5, alpha=.2) +
  geom_vline(
    aes(xintercept=mean(MDS_Scores_EMA_SI), color="mean"),
    color="#90a955", linetype="dashed", size=.8)+ #South
  geom_vline(
    aes(xintercept=mean(MDS_Scores_EMA_NI)),
    color="#212529", linetype="dashed", size=.8)+ #North
  scale_fill_manual(values=c("#212529","#90a955"))+
  scale_color_manual(values=c("#212529","#90a955"))+
  theme_pubclean()+
    labs(
      title="EMA",
      caption = paste0(
      "PERMANOVA: p < 0.0001 \n",
      "Wasserstein distance: ", round(
      wasserstein1d(MDS_Scores_EMA_SI,MDS_Scores_EMA_NI),2
      )))+
  theme(legend.position = "right")


mds_R.boxplot <- ggplot(mds_df_R, aes(x=Macroregion, y=MDS1, fill=Macroregion)) + 
    geom_boxplot(alpha=.2)+
  scale_fill_manual(values=c("#212529","#90a955"))+
  scale_color_manual(values=c("#212529","#90a955"))+
  theme_pubclean()+
    labs(
      title="Roman Age")+
  theme(legend.position = "none")

mds_EMA.boxplot <- ggplot(mds_df_EMA, aes(x=Macroregion, y=MDS1, fill=Macroregion)) + 
    geom_boxplot(alpha=.2)+
  scale_fill_manual(values=c("#212529","#90a955"))+
  scale_color_manual(values=c("#212529","#90a955"))+
  theme_pubclean()+
    labs(
      title="EMA")+
  theme(legend.position = "right")

```

```{r}
#| label: fig-nMDS-1D-macroregion
#| fig-cap: "nMDS reduction to one dimension performed on the Roman and early Medieval cereal datasets. The dashed colored lines indicate the mean of the distribution."
#| fig-subcap: 
#|  - "Roman age."
#|  - "Early Medieval Age."
#| layout-nrow: 1
#| layout-ncol: 2
#| echo: false


# Plot the two plots side by side
mds_R.plot 
mds_EMA.plot
```

```{r}
#| echo: false
#| label: fig-boxplot-nMDS-R-EMA
#| fig-cap: "Boxplots showing the nMDS scores for Northern and Southern Roman Italy."
#| fig-subcap: 
#|  - "Roman age."
#|  - "Early Medieval age."
#| layout-nrow: 1
#| layout-ncol: 2

mds_R.boxplot
mds_EMA.boxplot

```
:::

### Noble grains

```{r}
#| echo: false

# Convert to list
Noble_Grains.list <- list(
    NG_Found = Noble_Grains_Reg$Grain_Count,
    MG_Tot = 2, # Barley + Common Wheat
    RC_ID = as.factor((interaction(Noble_Grains_Reg$Chronology, 
                                   Noble_Grains_Reg$Macroregion)))
   )
```

```{r}
#| echo: false
#| eval: false
#| code-fold: true


# Create a binomial model, where N is 2 because that is the total number of noble 
# grains studied in this research
m_ng_reg <- ulam(
    alist(
        NG_Found ~ dbinom( 2 , p ),
        logit(p) <- RegChr[RC_ID],
        RegChr[RC_ID] ~ dnorm(0,1)
    ), data=Noble_Grains.list , chains=4
    )

#saveRDS(m_ng_reg, "stan_models/m_ng_reg.rds")

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| output: false


m_ng_reg <- readRDS("stan_models/m_ng_reg.rds")

m_ng_reg.df <- m_ng_reg %>%
  spread_draws(RegChr[RC_ID])

m_ng_reg.df %>% mutate(Region = case_when(
  RC_ID >= 9 &  RC_ID <=12 ~ "SI",
  RC_ID >= 1 & RC_ID <= 4 ~ "CI",
  RC_ID >= 5 & RC_ID <= 8  ~ "NI"
))  %>% mutate(Chrono = case_when(
    RC_ID == 4 |  RC_ID == 8 | RC_ID==12  ~ "R",
    RC_ID == 2 |  RC_ID == 6 | RC_ID == 10  ~ "LR",
    RC_ID == 1 |  RC_ID == 5 | RC_ID == 9  ~ "EMA",
    RC_ID == 3 |  RC_ID == 7 | RC_ID == 11 ~ "Ma"
)) -> m_ng_reg.df

m_ng_reg.plot <- ggplot(
  data = m_ng_reg.df,
  aes(
    y = (factor(Chrono, levels = c("R", "LR", "EMA", "Ma"))),
    x = inv_logit(RegChr)
  )
) +
  stat_interval(
    .width = c(.50, .80, .95, .99),
    linewidth = 3
  ) +
  geom_vline(
    xintercept = inv_logit(mean(m_ng_reg.df$RegChr)),
    linetype = "twodash",
    linewidth = 0.45,
    color = "red4"
  ) +
  scale_color_brewer(palette = "Greys") +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 10),
    sec.axis = dup_axis()
  ) +
  theme_tidybayes() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(
      fill = "grey",
      linewidth = 0.1,
      colour = "black"
    ),
    panel.background = element_rect(fill = "grey90"),
    panel.border = element_rect(fill = NA, linewidth = 0.1),
    panel.grid.major.x = element_line(
      color = "grey50",
      linewidth = 0.25,
      linetype = 2
    ),
    axis.title.x.top = element_blank()
  ) +
  facet_grid(factor(Region, levels=c("SI", "CI", "NI")) ~ ., scales = "free") +
  labs(
    title = "Noble Grains",
    subtitle = "Macroregional distribution",
    x = "Probability",
    y = "Chronology",
    color = "Credible interval",
    caption = "The dashed line indicates the mean for each macroregion and chronology."
  )

```

```{r}
#| echo: false
#| label: fig-bayes-glm-ng-macroregion
#| fig-cap: "Bayesian probability estimates for noble grains (common wheat and barley), divided by macroregion and chronology. The prediction estimates are produced with a GLM and weakly informative priors. Credible intervals are provided."

m_ng_reg.plot
```

### Minor grains

```{r}
#| echo: false
#| eval: true
#| code-fold: true

# Convert to list
Minor_Grains.list <- list(
    MG_Found = Minor_Grains$Grain_Count,
    MG_Tot = 7,
    RC_ID = as.factor((interaction(Minor_Grains$Chronology, 
                                   Minor_Grains$Macroregion)))
   )
```

```{r}
#| echo: false
#| eval: false
#| code-fold: true

# Create a binomial model, where N is 7 because that is the total number of minor 
# grains studied in this thesis
m_mg_reg <- ulam(
    alist(
        MG_Found ~ dbinom( 7 , p ),
        logit(p) <- RegChr[RC_ID],
        RegChr[RC_ID] ~ dnorm(0,1)
    ), data=Minor_Grains.list , chains=4
    )

#saveRDS(m_mg_reg, "stan_models/m_mg_reg.rds")

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| output: false

m_mg_reg <- readRDS("stan_models/m_mg_reg.rds")

m_mg_reg.df <- m_mg_reg %>%
  spread_draws(RegChr[RC_ID])

m_mg_reg.df %>% mutate(Region = case_when(
  RC_ID >= 9 &  RC_ID <=12 ~ "SI",
  RC_ID >= 1 & RC_ID <= 4 ~ "CI",
  RC_ID >= 5 & RC_ID <= 8  ~ "NI"
))  %>% mutate(Chrono = case_when(
    RC_ID == 4 |  RC_ID == 8 | RC_ID==12  ~ "R",
    RC_ID == 2 |  RC_ID == 6 | RC_ID == 10  ~ "LR",
    RC_ID == 1 |  RC_ID == 5 | RC_ID == 9  ~ "EMA",
    RC_ID == 3 |  RC_ID == 7 | RC_ID == 11 ~ "Ma"
)) -> m_mg_reg.df

m_mg_reg.plot <- ggplot(
  data = m_mg_reg.df,
  aes(
    y = (factor(Chrono, levels = c("R", "LR", "EMA", "Ma"))),
    x = inv_logit(RegChr)
  )
) +
  stat_interval(
    .width = c(.50, .80, .95, .99),
    linewidth = 3
  ) +
  geom_vline(
    xintercept = inv_logit(mean(m_mg_reg.df$RegChr)),
    linetype = "twodash",
    linewidth = 0.45,
    color = "red4"
  ) +
  scale_color_brewer(palette = "Greys") +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 10),
    sec.axis = dup_axis()
  ) +
  theme_tidybayes() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(
      fill = "grey",
      linewidth = 0.1,
      colour = "black"
    ),
    panel.background = element_rect(fill = "grey90"),
    panel.border = element_rect(fill = NA, linewidth = 0.1),
    panel.grid.major.x = element_line(
      color = "grey50",
      linewidth = 0.25,
      linetype = 2
    ),
    axis.title.x.top = element_blank()
  ) +
  facet_grid(factor(Region, levels=c("SI", "CI", "NI")) ~ ., scales = "free") +
  labs(
    title = "Minor Grains",
    subtitle = "Macroregional distribution",
    x = "Probability",
    y = "Chronology",
    color = "Credible interval",
    caption = "The dashed line indicates the mean for each macroregion and chronology."
  )

```

```{r}
#| echo: false
#| label: fig-bayes-glm-mg-macroregion
#| fig-cap: "Bayesian probability estimates for minor grains (Emmer, Einkorn, Foxtail and Broomcorn millets, Sorghum), divided by macroregion and chronology. The prediction estimates are produced with a GLM and weakly informative priors. Credible intervals are provided."

m_mg_reg.plot
```

## Legumes

```{r}
#| echo: false


####################
## LEGUMES 
####################

# Convert to list
Legumes_reg.list <- list(
    Leg_Found = Legumes$Seed_Count,
    Leg_Tot = 7,
    RC_ID = as.factor((interaction(Legumes$Chronology, 
                                   Legumes$Macroregion)))
   )

Legumes_reg.list$RC_ID <- droplevels(Legumes_reg.list$RC_ID)


```

```{r}
#| echo: false
#| eval: false
#| code-fold: true

# Create a binomial model, where N is 7 because that is the total number of minor 
# grains studied in this thesis
m_leg_reg <- ulam(
    alist(
        Leg_Found ~ dbinom( 7 , p ),
        logit(p) <- RegChr[RC_ID],
        RegChr[RC_ID] ~ dnorm(0,1)
    ), data=Legumes_reg.list , chains=4
    )

#saveRDS(m_leg_reg, "stan_models/m_leg_reg.rds")

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| output: false

m_leg_reg <- readRDS("stan_models/m_leg_reg.rds")

m_leg_reg.df <- m_leg_reg %>%
  spread_draws(RegChr[RC_ID])

m_leg_reg.df %>% mutate(Region = case_when(
  RC_ID >= 9 &  RC_ID <=12 ~ "SI",
  RC_ID >= 1 & RC_ID <= 4 ~ "CI",
  RC_ID >= 5 & RC_ID <= 8  ~ "NI"
))  %>% mutate(Chrono = case_when(
    RC_ID == 4 |  RC_ID == 8 | RC_ID==12  ~ "R",
    RC_ID == 2 |  RC_ID == 6 | RC_ID == 10  ~ "LR",
    RC_ID == 1 |  RC_ID == 5 | RC_ID == 9  ~ "EMA",
    RC_ID == 3 |  RC_ID == 7 | RC_ID == 11 ~ "Ma"
)) -> m_leg_reg.df

m_leg_reg.plot <- ggplot(
  data = m_leg_reg.df,
  aes(
    y = (factor(Chrono, levels = c("R", "LR", "EMA", "Ma"))),
    x = inv_logit(RegChr)
  )
) +
  stat_interval(
    .width = c(.50, .80, .95, .99),
    linewidth = 3
  ) +
  geom_vline(
    xintercept = inv_logit(mean(m_leg_reg.df$RegChr)),
    linetype = "twodash",
    linewidth = 0.45,
    color = "red4"
  ) +
  scale_color_brewer(palette = "Greys") +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 10),
    sec.axis = dup_axis()
  ) +
  theme_tidybayes() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(
      fill = "grey",
      linewidth = 0.1,
      colour = "black"
    ),
    panel.background = element_rect(fill = "grey90"),
    panel.border = element_rect(fill = NA, linewidth = 0.1),
    panel.grid.major.x = element_line(
      color = "grey50",
      linewidth = 0.25,
      linetype = 2
    ),
    axis.title.x.top = element_blank()
  ) +
  facet_grid(factor(Region, levels=c("SI", "CI", "NI")) ~ ., scales = "free") +
  labs(
    title = "Legumes",
    subtitle = "Macroregional distribution",
    x = "Probability",
    y = "Chronology",
    color = "Credible interval",
    caption = "The dashed line indicates the mean for each macroregion and chronology."
  )

```

```{r}
#| echo: false
#| label: fig-bayes-glm-legumes-macroregion
#| fig-cap: "Bayesian probability estimates for legumes, divided by macroregion and chronology. The prediction estimates are produced with a GLM and weakly informative priors. Credible intervals are provided."

m_leg_reg.plot

```

## Fruits

### Grape

```{r}
#| echo: false

Grape_reg.list <- list(
    Grape_Found = Grape$Present,
    RC_ID = as.factor((interaction(Legumes$Chronology, 
                                   Legumes$Macroregion)))
   )

Grape_reg.list$RC_ID <- droplevels(Grape_reg.list$RC_ID)
```

```{r}
#| echo: false
#| eval: false

m_grape_reg <- ulam(
    alist(
        Grape_Found ~ dbinom( 1 , p ),
        logit(p) <- RegChr[RC_ID],
        RegChr[RC_ID] ~ dnorm(0,1)
    ), data=Grape_reg.list , chains=4
    )

#saveRDS(m_grape_reg, "stan_models/m_grape_reg.rds")


```

```{r}
#| echo: false
#| warning: false
#| message: false
#| output: false

m_grape_reg <- readRDS("stan_models/m_grape_reg.rds")

m_grape_reg.df <- m_grape_reg %>%
  spread_draws(RegChr[RC_ID])

m_grape_reg.df %>% mutate(Region = case_when(
  RC_ID >= 9 &  RC_ID <=12 ~ "SI",
  RC_ID >= 1 & RC_ID <= 4 ~ "CI",
  RC_ID >= 5 & RC_ID <= 8  ~ "NI"
))  %>% mutate(Chrono = case_when(
    RC_ID == 4 |  RC_ID == 8 | RC_ID==12  ~ "R",
    RC_ID == 2 |  RC_ID == 6 | RC_ID == 10  ~ "LR",
    RC_ID == 1 |  RC_ID == 5 | RC_ID == 9  ~ "EMA",
    RC_ID == 3 |  RC_ID == 7 | RC_ID == 11 ~ "Ma"
)) -> m_grape_reg.df

m_grape_reg.plot <- ggplot(
  data = m_grape_reg.df,
  aes(
    y = (factor(Chrono, levels = c("R", "LR", "EMA", "Ma"))),
    x = inv_logit(RegChr)
  )
) +
  stat_interval(
    .width = c(.50, .80, .95, .99),
    linewidth = 3
  ) +
  geom_vline(
    xintercept = inv_logit(mean(m_grape_reg.df$RegChr)),
    linetype = "twodash",
    linewidth = 0.45,
    color = "red4"
  ) +
  scale_color_brewer(palette = "Greys") +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 10),
    sec.axis = dup_axis()
  ) +
  theme_tidybayes() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(
      fill = "grey",
      linewidth = 0.1,
      colour = "black"
    ),
    panel.background = element_rect(fill = "grey90"),
    panel.border = element_rect(fill = NA, linewidth = 0.1),
    panel.grid.major.x = element_line(
      color = "grey50",
      linewidth = 0.25,
      linetype = 2
    ),
    axis.title.x.top = element_blank()
  ) +
  facet_grid(factor(Region, levels=c("SI", "CI", "NI")) ~ ., scales = "free") +
  labs(
    title = "Grape",
    subtitle = "Macroregional distribution",
    x = "Probability",
    y = "Chronology",
    color = "Credible interval",
    caption = "The dashed line indicates the mean for each macroregion and chronology."
  )

```

```{r}
#| echo: false
#| label: fig-bayes-glm-grape-macroregion
#| fig-cap: "Bayesian probability estimates for grape, divided by macroregion and chronology. The prediction estimates are produced with a GLM and weakly informative priors. Credible intervals are provided."

m_grape_reg.plot

```

### Olives

```{r}
#| echo: false


# Convert to list
Olive_reg.list <- list(
    Olive_Found = Olives$Present,
    RC_ID = as.factor((interaction(Olives$Chronology, 
                                   Olives$Macroregion)))
   )

Olive_reg.list$RC_ID <- droplevels(Olive_reg.list$RC_ID)
  
```

```{r}
#| echo: false
#| eval: false

m_olive_reg <- ulam(
    alist(
        Olive_Found ~ dbinom( 1 , p ),
        logit(p) <- RegChr[RC_ID],
        RegChr[RC_ID] ~ dnorm(0,1)
    ), data=Olive_reg.list , chains=4
    )

saveRDS(m_olive_reg, "stan_models/m_olive_reg.rds")

```

```{r}
#| echo: false
#| warning: false
#| message: false
#| output: false

m_olive_reg <- readRDS("stan_models/m_olive_reg.rds")

m_olive_reg.df <- m_olive_reg %>%
  spread_draws(RegChr[RC_ID])

m_olive_reg.df %>% mutate(Region = case_when(
  RC_ID >= 9 &  RC_ID <=12 ~ "SI",
  RC_ID >= 1 & RC_ID <= 4 ~ "CI",
  RC_ID >= 5 & RC_ID <= 8  ~ "NI"
))  %>% mutate(Chrono = case_when(
    RC_ID == 4 |  RC_ID == 8 | RC_ID==12  ~ "R",
    RC_ID == 2 |  RC_ID == 6 | RC_ID == 10  ~ "LR",
    RC_ID == 1 |  RC_ID == 5 | RC_ID == 9  ~ "EMA",
    RC_ID == 3 |  RC_ID == 7 | RC_ID == 11 ~ "Ma"
)) -> m_olive_reg.df

m_olive_reg.plot <- ggplot(
  data = m_olive_reg.df,
  aes(
    y = (factor(Chrono, levels = c("R", "LR", "EMA", "Ma"))),
    x = inv_logit(RegChr)
  )
) +
  stat_interval(
    .width = c(.50, .80, .95, .99),
    linewidth = 3
  ) +
  geom_vline(
    xintercept = inv_logit(mean(m_olive_reg.df$RegChr)),
    linetype = "twodash",
    linewidth = 0.45,
    color = "red4"
  ) +
  scale_color_brewer(palette = "Greys") +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 10),
    sec.axis = dup_axis()
  ) +
  theme_tidybayes() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(
      fill = "grey",
      linewidth = 0.1,
      colour = "black"
    ),
    panel.background = element_rect(fill = "grey90"),
    panel.border = element_rect(fill = NA, linewidth = 0.1),
    panel.grid.major.x = element_line(
      color = "grey50",
      linewidth = 0.25,
      linetype = 2
    ),
    axis.title.x.top = element_blank()
  ) +
  facet_grid(factor(Region, levels=c("SI", "CI", "NI")) ~ ., scales = "free") +
  labs(
    title = "Olive",
    subtitle = "Macroregional distribution",
    x = "Probability",
    y = "Chronology",
    color = "Credible interval",
    caption = "The dashed line indicates the mean for each macroregion and chronology."
  )

```

```{r}
#| echo: false
#| label: fig-bayes-glm-olive-macroregion
#| fig-cap: "Bayesian probability estimates for olive, divided by macroregion and chronology. The prediction estimates are produced with a GLM and weakly informative priors. Credible intervals are provided."

m_olive_reg.plot

```

### Nuts

```{r}
#| echo: false

####################
## NUTS
####################

# Convert to list
Nuts_reg.list <- list(
    Nuts_Found = Nuts$Seed_Count,
    Nuts_Tot = 3,
    RC_ID = as.factor((interaction(Nuts$Chronology, 
                                   Nuts$Macroregion)))
   )

Nuts_reg.list$RC_ID <- droplevels(Nuts_reg.list$RC_ID)

```

```{r}
#| echo: false
#| eval: false

m_nuts_reg <- ulam(
    alist(
        Nuts_Found ~ dbinom( 3 , p ),
        logit(p) <- RegChr[RC_ID],
        RegChr[RC_ID] ~ dnorm(0,1)
    ), data=Nuts_reg.list , chains=4
    )

saveRDS(m_nuts_reg, "stan_models/m_nuts_reg.rds")


```

```{r}
#| echo: false
#| warning: false
#| message: false
#| output: false

m_nuts_reg <- readRDS("stan_models/m_nuts_reg.rds")

m_nuts_reg.df <- m_nuts_reg %>%
  spread_draws(RegChr[RC_ID])

m_nuts_reg.df %>% mutate(Region = case_when(
  RC_ID >= 9 &  RC_ID <=12 ~ "SI",
  RC_ID >= 1 & RC_ID <= 4 ~ "CI",
  RC_ID >= 5 & RC_ID <= 8  ~ "NI"
))  %>% mutate(Chrono = case_when(
    RC_ID == 4 |  RC_ID == 8 | RC_ID==12  ~ "R",
    RC_ID == 2 |  RC_ID == 6 | RC_ID == 10  ~ "LR",
    RC_ID == 1 |  RC_ID == 5 | RC_ID == 9  ~ "EMA",
    RC_ID == 3 |  RC_ID == 7 | RC_ID == 11 ~ "Ma"
)) -> m_nuts_reg.df

m_nuts_reg.plot <- ggplot(
  data = m_nuts_reg.df,
  aes(
    y = (factor(Chrono, levels = c("R", "LR", "EMA", "Ma"))),
    x = inv_logit(RegChr)
  )
) +
  stat_interval(
    .width = c(.50, .80, .95, .99),
    linewidth = 3
  ) +
  geom_vline(
    xintercept = inv_logit(mean(m_nuts_reg.df$RegChr)),
    linetype = "twodash",
    linewidth = 0.45,
    color = "red4"
  ) +
  scale_color_brewer(palette = "Greys") +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 10),
    sec.axis = dup_axis()
  ) +
  theme_tidybayes() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(
      fill = "grey",
      linewidth = 0.1,
      colour = "black"
    ),
    panel.background = element_rect(fill = "grey90"),
    panel.border = element_rect(fill = NA, linewidth = 0.1),
    panel.grid.major.x = element_line(
      color = "grey50",
      linewidth = 0.25,
      linetype = 2
    ),
    axis.title.x.top = element_blank()
  ) +
  facet_grid(factor(Region, levels=c("SI", "CI", "NI")) ~ ., scales = "free") +
  labs(
    title = "Nuts",
    subtitle = "Macroregional distribution",
    x = "Probability",
    y = "Chronology",
    color = "Credible interval",
    caption = "The dashed line indicates the mean for each macroregion and chronology."
  )

```

```{r}
#| echo: false
#| label: fig-bayes-glm-nuts-macroregion
#| fig-cap: "Bayesian probability estimates for nuts, divided by macroregion and chronology. The prediction estimates are produced with a GLM and weakly informative priors. Credible intervals are provided."

m_nuts_reg.plot

```

### Domestic and wild fruits

```{r}
#| echo: false

####################
## Domestic FRUITS 
####################

# Convert to list
DomFruit_reg.list <- list(
    DF_Found = DomFruit$Seed_Count,
    DF_Tot = 7,
    RC_ID = as.factor((interaction(DomFruit$Chronology, 
                                   DomFruit$Macroregion)))
   )

DomFruit_reg.list$RC_ID <- droplevels(DomFruit_reg.list$RC_ID)

```

```{r}
#| echo: false
#| eval: false

m_dfruit_reg <- ulam(
    alist(
        DF_Found ~ dbinom( 7 , p ),
        logit(p) <- RegChr[RC_ID],
        RegChr[RC_ID] ~ dnorm(0,1)
    ), data=DomFruit_reg.list , chains=4
    )

m_berries_reg <- ulam(
    alist(
        Berries_Found ~ dbinom( 9 , p ),
        logit(p) <- RegChr[RC_ID],
        RegChr[RC_ID] ~ dnorm(0,1)
    ), data=Berries_reg.list , chains=4
    )

#saveRDS(m_dfruit_reg, "stan_models/m_dfruit_reg.rds")
#saveRDS(m_berries_reg, "stan_models/m_berries_reg.rds")


```

```{r}
#| echo: false
#| warning: false
#| message: false
#| output: false

m_berries_reg <- readRDS("stan_models/m_berries_reg.rds")

m_berries_reg.df <- m_berries_reg %>%
  spread_draws(RegChr[RC_ID])

m_berries_reg.df %>% mutate(Region = case_when(
  RC_ID >= 9 &  RC_ID <=12 ~ "SI",
  RC_ID >= 1 & RC_ID <= 4 ~ "CI",
  RC_ID >= 5 & RC_ID <= 8  ~ "NI"
))  %>% mutate(Chrono = case_when(
  RC_ID == 4 |  RC_ID == 8 | RC_ID==12  ~ "R",
  RC_ID == 2 |  RC_ID == 6 | RC_ID == 10  ~ "LR",
  RC_ID == 1 |  RC_ID == 5 | RC_ID == 9  ~ "EMA",
  RC_ID == 3 |  RC_ID == 7 | RC_ID == 11 ~ "Ma"
)) -> m_berries_reg.df

m_berries_reg.plot <- ggplot(
  data = m_berries_reg.df,
  aes(
    y = (factor(Chrono, levels = c("R", "LR", "EMA", "Ma"))),
    x = inv_logit(RegChr)
  )
) +
  stat_interval(
    .width = c(.50, .80, .95, .99),
    linewidth = 3
  ) +
  geom_vline(
    xintercept = inv_logit(mean(m_berries_reg.df$RegChr)),
    linetype = "twodash",
    linewidth = 0.45,
    color = "red4"
  ) +
  scale_color_brewer(palette = "Greys") +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 10),
    sec.axis = dup_axis()
  ) +
  theme_tidybayes() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(
      fill = "grey",
      linewidth = 0.1,
      colour = "black"
    ),
    panel.background = element_rect(fill = "grey90"),
    panel.border = element_rect(fill = NA, linewidth = 0.1),
    panel.grid.major.x = element_line(
      color = "grey50",
      linewidth = 0.25,
      linetype = 2
    ),
    axis.title.x.top = element_blank()
  ) +
  facet_grid(factor(Region, levels=c("SI", "CI", "NI")) ~ ., scales = "free") +
  labs(
    title = "Berries",
    subtitle = "Macroregional distribution",
    x = "Probability",
    y = "Chronology",
    color = "Credible interval",
    caption = "The dashed line indicates the mean for each macroregion and chronology."
  )


m_dfruit_reg <- readRDS("stan_models/m_dfruit_reg.rds")

m_dfruit_reg.df <- m_dfruit_reg %>%
  spread_draws(RegChr[RC_ID])

m_dfruit_reg.df %>% mutate(Region = case_when(
  RC_ID >= 9 &  RC_ID <=12 ~ "SI",
  RC_ID >= 1 & RC_ID <= 4 ~ "CI",
  RC_ID >= 5 & RC_ID <= 8  ~ "NI"
))  %>% mutate(Chrono = case_when(
  RC_ID == 4 |  RC_ID == 8 | RC_ID==12  ~ "R",
  RC_ID == 2 |  RC_ID == 6 | RC_ID == 10  ~ "LR",
  RC_ID == 1 |  RC_ID == 5 | RC_ID == 9  ~ "EMA",
  RC_ID == 3 |  RC_ID == 7 | RC_ID == 11 ~ "Ma"
)) -> m_dfruit_reg.df

m_dfruit_reg.plot <- ggplot(
  data = m_dfruit_reg.df,
  aes(
    y = (factor(Chrono, levels = c("R", "LR", "EMA", "Ma"))),
    x = inv_logit(RegChr)
  )
) +
  stat_interval(
    .width = c(.50, .80, .95, .99),
    linewidth = 3
  ) +
  geom_vline(
    xintercept = inv_logit(mean(m_dfruit_reg.df$RegChr)),
    linetype = "twodash",
    linewidth = 0.45,
    color = "red4"
  ) +
  scale_color_brewer(palette = "Greys") +
  scale_y_discrete(limits = rev) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 10),
    sec.axis = dup_axis()
  ) +
  theme_tidybayes() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(
      fill = "grey",
      linewidth = 0.1,
      colour = "black"
    ),
    panel.background = element_rect(fill = "grey90"),
    panel.border = element_rect(fill = NA, linewidth = 0.1),
    panel.grid.major.x = element_line(
      color = "grey50",
      linewidth = 0.25,
      linetype = 2
    ),
    axis.title.x.top = element_blank()
  ) +
  facet_grid(factor(Region, levels=c("SI", "CI", "NI")) ~ ., scales = "free") +
  labs(
    title = "Domestic Fruits",
    subtitle = "Macroregional distribution",
    x = "Probability",
    y = "Chronology",
    color = "Credible interval",
    caption = "The dashed line indicates the mean for each macroregion and chronology."
  )

```

```{r}
#| echo: false
#| label: fig-bayes-glm-domfruits-macroregion
#| fig-cap: "Bayesian probability estimates for domestic fruits, divided by macroregion and chronology. The prediction estimates are produced with a GLM and weakly informative priors. Credible intervals are provided."

m_dfruit_reg.plot

```

```{r}
#| echo: false
#| label: fig-bayes-glm-berries-macroregion
#| fig-cap: "Bayesian probability estimates for domestic fruits, divided by macroregion and chronology. The prediction estimates are produced with a GLM and weakly informative priors. Credible intervals are provided."

m_berries_reg.plot

```

# To do:

## Wine/Olive oil production in Italy

Questions: Is there a decrease in the production of cash crops in Late Roman villas? In particular in central Italian villas.

Available data:

-   Seed remains

-   Wine presses in Roman Italy (1st to 6th c. ca) -\> Dodd

-   Wine/Olive presses, dated but no info on wine/olive or size: http://oxrep.classics.ox.ac.uk/databases/olive_oil_and_wine_presses_database/

-   Dodd: "a notable increase in size from around the first century BCE." No quantitative data provided.

## Fruit consumption

-   The domesticated and the wild

-   What fruits arrive on the Roman table?

-   Can we see a decreased consumption of figs during the LALIA in the north?

::: callout-alert
This kind of data manipulation might be misleading as it is ignoring the fact that some centuries are only represented by - let's say - two sites. Is this a good idea?
:::

```{r}
#| echo: false
#| warning: false
#| message: false

######################
## Fig presence
######################

  figs <- with(
  Df_Cond_Plants,
data.frame(
  Site, name_macroreg, Chronology, From.Century, To.Century, 
  Present=decostand(Df_Cond_Plants[11:ncol(Df_Cond_Plants)], method="pa")$Fig
  )
)

# North
figs.North <- subset(figs, name_macroreg=="Northern Italy")
figs.North$From.Century <- figs.North$From.Century*100
figs.North$To.Century <- figs.North$To.Century*100
figs.North <- figs.North[!duplicated(figs.North[-3]), ]

figs.North <- figs.North %>% 
    mutate(year = map2(From.Century, To.Century, `:`)) %>% 
    select(-From.Century, -To.Century) %>% 
    unnest

# South
figs.South <- subset(figs, name_macroreg=="Southern Italy")
figs.South$From.Century <- figs.South$From.Century*100
figs.South$To.Century <- figs.South$To.Century*100
figs.South <- figs.South[!duplicated(figs.South[-3]), ]

figs.South <- figs.South %>% 
    mutate(year = map2(From.Century, To.Century, `:`)) %>% 
    select(-From.Century, -To.Century) %>% 
    unnest


# Central
figs.Central <- subset(figs, name_macroreg=="Central Italy")
figs.Central$From.Century <- figs.Central$From.Century*100
figs.Central$To.Century <- figs.Central$To.Century*100
figs.Central <- figs.Central[!duplicated(figs.Central[-3]), ]

figs.Central <- figs.Central %>% 
    mutate(year = map2(From.Century, To.Century, `:`)) %>% 
    select(-From.Century, -To.Century) %>% 
    unnest
```

```{r}
#| echo: false
#| fig-height: 2
#| label: fig-figs-cdplots
#| fig-cap: "The values for Central Italy and for Late Roman Southern Italy are based on very few contexts."


par(mfrow=c(1,3))
cdplot(figs.South$year, as.factor(figs.South$Present), 
       xlim=c(-100,1100),
       main="Presence of figs in S Italy",
       xlab="Year",
       ylab="Presence"
       )

cdplot(figs.Central$year, as.factor(figs.Central$Present), 
       xlim=c(-100,1100),
       main="Presence of figs in C Italy",
       sub="Very few archaeobotanical contexts.",
       xlab="Year",
       ylab="Presence"
       )

cdplot(figs.North$year, as.factor(figs.North$Present), 
       xlim=c(-100,1100),
       main="Presence of figs in N Italy",
       xlab="Year",
       ylab="Presence"
       )


```
